---
title: "Assignment 1"
output: 
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(sf)
library(tidyverse)
library(here)
library(rnaturalearth)

countries <- ne_countries(continent = "Africa",
                          returnclass = "sf")

airports <- here("a1-files",
              "sample-files",
              "airports") %>%
  st_read() %>%
  st_filter(countries)

rail <- here("a1-files",
              "sample-files",
              "rail") %>%
  st_read() %>%
  st_filter(countries)
```

I have selected the following three datasets, all of which are in the 
WGS 84 coordinate reference system:

1. A set of polygons representing national boundaries in Africa (shown below 
as gray polygons with black borders). They are downloaded from Natural Earth 
Data via the rnaturalearth package as R simple features. The dataset 
includes 51 polygons, each representing a country. There are 64 variables, 
including the following that I might include in future analyses:
    * pop_est: The estimated population
    * gdp_md_est: The estimated GDP in millions of US dollars 
    * economy: Economic classification as "Emerging region: G20", "Developing region", or "Least developed region".
    * income_grp: Classification as high income, upper middle income, 
    lower middle income, or low income.
2. A set of lines representing railway lines in Africa, shown below as red 
lines. I downloaded them from Natural Earth Data (www.naturalearthdata.com) 
via the web interface as an ESRI shapefile including global railways, then I 
filtered them to include only the railways located within the set of polygons
described above. There are 751 rail lines included in the data and 13 
variables. I don't anticipate using any of these 13 variables in future 
analysis.
3. A set of points representing airports in Africa, shown below as black 
dots. I downloaded them from Natural Earth Data (www.naturalearthdata.com) 
via the web interface as an ESRI shapefile including global airports, then I 
filtered them to include only points located within the set of polygons 
representing African countries. There are 85 rail lines included in the data 
and 41 variables, 27 of which are the names of the airport in each of 27 
languages. I don't anticipate using any of these variables in future 
analysis.

```{r, echo=FALSE}
ggplot(countries) +
  geom_sf() +
  geom_sf(data = airports) +
  geom_sf(data = rail, color = "red")
```

My classmate, *Jonathan Lee*, helped me complete this assignment by brainstorming
with me about potential topics.
