---
title: "Assignment 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Generate a map to communicate the travel times between features

* using isochrones _*and*_ 
* by communicating continuous variations in travel time  
    * by symbolizing features based on their travel time (by any mode) from other features _*or*_
    * showing descriptive statistics such as averages and standard deviations _*or*_
    * illustrating distributions and relationships using charts and graphs

Your map layout should 
include a description of the analysis question your
map is meant to address and a clear description of what each 
layer represents (this might be in the map legend). 

Include the sources of all the data shown
in your map (including any data that isn't shown but was 
used to derive the layers you've shown), the name of a
classmate who was helpful to you, and how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this 
assignment, but all of them are relevant and you will 
probably use some of them. 

#### R

These do build on one another.

* [Generating a travel time matrix](){target="_blank"} (this is a necessary preliminary step for finding the closest point and for generating isochrones)
* [Finding the travel time to the closest point](){target="_blank"}
* [Generating isochrones](){target="_blank"}
* [Calculating average values for a variable](){target="_blank"}
* [Non-map charts and graphs](){target="_blank"}

#### ArcGIS Pro

* [Finding the travel time to the closest point]



# Examples

The examples below fully meet the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/Examples/slo-trails.html){target="_blank"}
* [Example 2](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/Examples/tornado-example.pdf){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Inclusion of isochrones",
                          "Illustration of continuous travel time",
                          "Description of analysis question",
                          "Description of conclusion",
                          "Source attribution",
                          "Total"),
                 `No credit` = c(rep("Missing", 7), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "",
                                   "",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described",
                                   ""),
                 `Full credit` = c("Adequately described",
                                   "Adequately described",
                                   "Included",
                                   "Included",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 5, 5, 5, 5, 1, 25))

kable(rubric)
```
