---
title: "Simple charts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A map is one of many ways you can visualize spatial data.
Scatter plots and and histograms can be good alternatives.

# Required packages

This example uses the following packages. 

```{r, results='hide', message=FALSE}
library(tidyverse)
library(tidycensus)
library(knitr)
library(here)
```

# Sample data

I'll start by loading data on school districts in California. There are 
two data column: The proportion of the district population that are
children between the ages of 6 and 17, and the median income of district
households.

```{r, include=FALSE}
vars <- c(med_inc = "B06011_001",
          num_school_kids = "B05009_020",
          population = "B01003_001")

districts <- get_acs(geography = "school district (unified)",
                    state = "CA",
                    variables = vars,
                    output = "wide",
                    geometry = TRUE) %>%
  mutate(pct_kids = num_school_kidsE / populationE) %>%
  rename(med_income = med_incE) %>%
  select(med_income, pct_kids)
```

```{r}
head(districts) %>%
  kable()
```

# Scatter plot

I can create a scatter plot to show the relationship between these
two variables.

```{r, warning=FALSE}
ggplot(districts) +
  geom_point(aes(x = med_income,
                 y = pct_kids),
             size = 0.5) +
  scale_x_continuous(
    name = "Median income within school district boundaries",
    breaks = breaks <- seq(0, 100000, by = 20000),
    labels = paste0("$", 
                    formatC(breaks, 
                            big.mark = ",",
                            format = "f",
                            digits = 0))) +
  scale_y_continuous(
    name = "Percent of population between\n6 and 17 years old",
    breaks = breaks <- seq(0, 0.25, by=0.05),
    labels = paste0(breaks*100, "%")) +
  theme_minimal()
```

Scatter plots can also be useful for showing variation in a single 
variable. `position = "jitter"` can be used in this case to distribute
the points randomly across the vertical axis.

```{r, warning=FALSE}
ggplot(districts) +
  geom_point(aes(x = med_income,
                 y = 1),
             size = 0.5,
             position = "jitter") +
  scale_x_continuous(
    name = "Median income within school district boundaries",
    breaks = breaks <- seq(0, 100000, by = 20000),
    labels = paste0("$", 
                    formatC(breaks, 
                            big.mark = ",",
                            format = "f",
                            digits = 0))) +
  scale_y_continuous(name = "",
                     breaks = c(0,2)) +
  theme_minimal() +
  theme(axis.text.y = element_blank())
```


# Histograms

Histograms are a more traditional way to show variation in a 
single variable.

```{r}
ggplot(districts) +
  geom_histogram(aes(x = pct_kids),
                 binwidth = 0.01,
                 color = "lightgray",
                 fill = "darkgray") +
  scale_y_continuous(name = "Number of school districts") +
  scale_x_continuous(
    name = "Percent of population between\n6 and 17 years old",
    breaks = breaks <- seq(0, 0.25, by = 0.05),
    labels = paste0(breaks * 100, "%")) +
  theme_minimal()
```

# Save your chart

You can save your chart the same way you would save a map you 
created with ggplot.

```{r}
here("a5-files",
     "Examples",
     "histogram.png") %>%
  ggsave(height = 3, 
         width = 3, 
         units = "in",
         dpi = 300)
```

