---
title: "Calculating the average of a variable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It can sometimes be useful to summarize a variable in your
data by calculating an average.

# Required packages

This example uses the following packages. 

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(knitr)
```

# Sample data

```{r, include=FALSE}
library(here)
co_trails <- here("a1-files",
                  "County_Trails.geojson") %>%
  st_read() %>%
  select(OBJECTID, TrailName, Length)
```

In this example, have a dataset county-managed trails in San Luis 
Obispo County, California. There are three variables: an ID number,
the name of the trail, and the trail length (in miles).

```{r, warning=FALSE, message=FALSE}
head(co_trails) %>%
  kable()
```

# Calculate the average of a variable

The `mean()` function will return the average value for a variable.
Here's how I would calculate the average length of a trail.

```{r}
mean(co_trails$Length)
```
The average length of a trail is about 1.66 miles.