---
title: "Assignment 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Calculate the density of points (number of points per unit of area) or lines
(length of line per unit of area) within a layer of polygons. The polygon layer 
can be one that represents a set of features, or one that you have generated 
(for instance, a set of buffers or isochrones or a grid).

Display the results on a map. Include a statement about the question the
map is meant to answer and the answer to that question. You should also 
include the name of of a classmate who was helpful to you and a description 
of how they helped.

# Relevant Skills

#### R

* [Calculate the area of polygons within a polygon layer](https://c-voulgaris.github.io/VIS-2128-F22/a6-files/skill-calculate-area.html){target="_blank"}
* [Count the number of points within a polygon layer](https://c-voulgaris.github.io/VIS-2128-F22/a6-files/skill-count-points.html){target="_blank"}
* [Calculate the length of lines within a polygon layer](https://c-voulgaris.github.io/VIS-2128-F22/a6-files/skill-count-length.html){target="_blank"}

#### ArcGIS Pro

* [Calculate the areas of polygons](https://harvard.zoom.us/rec/share/_Na_H0miN8UTISITJrSu9ysDSv29bEGFG2jRkh-Ho1x0QW3QTKW-MLtp__It8cbG.hL733mCAFWuHXm_O?startTime=1665419672000){target="_blank"}
* [Count the number of points within a polygon layer](https://harvard.zoom.us/rec/share/LiPUEsKl2uGQXoAQvbraE9Tvkkyga7Rb2GShqhIj5kXnqGnYjCreCIulTA2EOVHU.j0Kqlo8hV2ga5Aqd?startTime=1665418912000){target="_blank"}
* [Calculate the length of lines within a polygon layer](https://harvard.zoom.us/rec/share/cbBv1vvzKLWZgt7dV8dYT-u5OgKdb0e3FP4GFXL7sDWFarTs9OnBK5-L7LrZi1wI.YdBPcc3H0iVFs7d7?startTime=1665422274000){target="_blank"}

# Additional resources

* [Lecture slides](https://c-voulgaris.github.io/VIS-2128-F22/a6-files/area-density-lecture.pdf){target="_blank"}

# Example

These examples would receive full credit on this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a6-files/Examples/park-map.html){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Description of density variable",
                          "Illustration of density variable",
                          "Description of analysis question",
                          "Description of conclusion",
                          "Source attribution",
                          "Total"),
                 `No credit` = c(rep("Missing", 7), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not clearly illustrated",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described",
                                   ""),
                 `Full credit` = c("Adequately described",
                                   "Adequately described",
                                   "Adequately described",
                                   "Clearly illustrated",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 5, 5, 5, 5, 1, 25))

kable(rubric)
```
