---
title: "Calculate the area of a polygon layer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Density is the ratio of the the amount of something within a 
polygon to the area of the polygon. This page demonstrates how 
to calculate the area of a polygon.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(knitr)
```

# Load some sample data

I'll start by downloading and loading a layer of polygons representing 
open space in Boston.

```{r}
parks <- st_read("https://bostonopendata-boston.opendata.arcgis.com/datasets/boston::open-space.geojson")
```

# Check for an area variable

The layer you are working with might already have a variable with the area 
of each polygon. Here are the variables included in our openspace data.

```{r}
colnames(parks)
```  

Two of these variables give the area of each polygon. `ACRES` is the area in 
(you guessed it) acres and `Shape_STArea` is the area is square feet. 

For purposes of this demonstration though, let's suppose you had to calculate the area.

# Transform the data to a projected coordinate system

Area calculations are based on coordinates, so you should transform your data to a 
projected coordinate system be for doing any calculations. I'll use the Massachusetts 
State Plane system, which has coordinates in units of feet. 

```{r}
MA_st_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +to_meter=0.3048006096012192 +no_defs" 

parks <- parks %>%
  st_transform(MA_st_plane)
```

# Calculate the area

`st_area()` will calculate the area of each polygon in the same units as its
coordinate reference system. You'll see that it matches the values for the 
`Shape_STArea` variable

```{r}
parks <- parks %>%
  mutate(area_feet_calc = st_area(.))

parks %>%
  select(ShapeSTArea, area_feet_calc) %>%
  head() %>%
  kable()
```

# Convert units to numbers

You'll notice that the units are labeled in the variable you just created. This is 
because it's special data type called units.

```{r}
class(parks$area_feet_calc)
```

This could trip you up if you want to do further calculations, so you might 
want to just convert it to a normal numeric variable.

```{r}
parks <- parks %>%
  mutate(area_feet_calc = as.numeric(area_feet_calc))

class(parks$area_feet_calc)
```

