---
title: "Count points within polygons"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Density is the ratio of the the amount of something within a 
polygon to the area of the polygon. This page demonstrates how 
to count the number of points within each polygon in a layer.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(knitr)
```

# Load some sample data

I'll start by downloading and loading a layer of polygons representing 
open space in Boston and a layer representing trees on public properties.

```{r}
MA_st_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +to_meter=0.3048006096012192 +no_defs" 

trees <- st_read("https://bostonopendata-boston.opendata.arcgis.com/datasets/boston::trees.geojson") %>%
  st_transform(MA_st_plane)

parks <- st_read("https://bostonopendata-boston.opendata.arcgis.com/datasets/boston::open-space.geojson") %>%
  st_transform(MA_st_plane)

```

# Count the points in each polygon

`st_intersects()` will return a list of indices for points within 
each polygon. We're really just interested in how many points are in 
each list, so I'll use `lengths()` to get the number of points in each 
polygon.

```{r}
parks <- parks %>%
  mutate(trees_in_park = lengths(st_intersects(., trees)))
```  

Now we have a variable called `trees_in_park` with the number of trees
within each park.

```{r}
parks %>%
  select(SITE_NAME, trees_in_park) %>%
  head() %>%
  kable()
```


