---
title: "Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Submit a description each of at least three spatial datasets with approximately 
the same spatial extent. At least one dataset should be a set of points; at least 
one should be a set of polygons; and at least one should be a set of lines. Your 
description can be entirely text, and it may include tables and other graphics. 
You should include the following information about each dataset:

* The spatial extent
* The unit of observation
* The number of observations
* The spatial data type (e.g. points, lines, or polygons)
* The spatial data format (e.g. KML, geoJSON, shapefile, R simple features)
* A description of each variable (that you might use in future analysis)
* The coordinate reference system
* The data source (who created the data and where did you find it)

Choose wisely and carefully! You will be working with these three 
datasets for the remainder of this course.

On your submission, list the name(s) of at least one student who was 
especially helpful you you in completing this assignment, and describe 
how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this assignment, but all 
of them are relevant and you will probably use several of them. They are
listed alphabetically.

* Aspatial filtering (ArcGIS) (link to tutorial forthcoming)
* [Aspatial filtering (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-aspatial-filter.html){target="_blank"}
* Converting between coordinate reference systems (ArcGIS) (link to tutorial forthcoming)
* [Converting between coordinate reference systems (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-convert-crs.html){target="_blank"}
* Determining the coordinate reference system for spatial data (ArcGIS) (link to tutorial forthcoming)
* [Determining the coordinate reference system for spatial data (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-get-crs.html){target="_blank"}
* [Getting national and subnational boundaries from rnaturalearth (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-rnaturalearth.html){target="_blank"}
* [Getting spatial data from a municipal open data portal](https://harvard.zoom.us/rec/play/J2QqcFKKvWv6MuPYGof2qnIYPXkc3ZJAVymlYITcFQ-t9IC4TDu-pFBoRcUTat3s0LtVx-qKNrSeHNcS.TOsORH6w8UnLGQRb?autoplay=true&startTime=1600374226000){target="_blank"}
* [Getting spatial data from the tidycensus package (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-tidycensus.html){target="_blank"}
* [Getting spatial data from the tigris package (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-tigris.html){target="_blank"}
* [Getting spatial data from U.S. Census TIGER files (web interface)](https://harvard.zoom.us/rec/play/QV2SCI71O-N_cgWLR0k0BMTb_WeldF-C4CaEhNxOcpJ7HcdM5nv3g1Lmv_nyqQ7XOboT2HWXhRkJblNA.Gd0LSXr6MkFzVIYC?startTime=1602525515000){target="_blank"}
* [Getting street data from OpenStreetMap (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-load-osm-data.html){target="_blank"}
* [Getting transit stop locations from a GTFS feed](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-get_gtfs-data.html){target="_blank"}
* Loading point data stored as coordinates in a csv file (ArcGIS) (link to tutorial forthcoming)
* Opening a shapefile (ArcGIS) (link to tutorial forthcoming)
* [Opening a spatial data set from a file (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-open-file.html){target="_blank"}
* [Plotting spatial data using ggplot (static) (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-ggplot-map.html){target="_blank"}
* Plotting spatial data using leaflet (interactive) (R)
* [Saving spatial data (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-save-spatial-data.html){target="_blank"}
* Spatial filtering (ArcGIS)  (link to tutorial forthcoming)
* [Spatial filtering (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-aspatial-filter.html){target="_blank"}
* Viewing spatial data as a table (ArcGIS) (link to tutorial forthcoming)
* [Viewing spatial data as a table (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-view-dataframe.html){target="_blank"}

# Other resources

* [The World Bank Data Catalog](https://datacatalog.worldbank.org/home){target="_blank"}
* [The Africa GeoPortal](https://www.africageoportal.com/){target="_blank"}
* [ESRI Open Data Hub](https://hub.arcgis.com/search){target="_blank"}
* [Natural Earth Data](https://www.naturalearthdata.com/){target="_blank"}

# Examples

Each of the examples below fully meets the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/examples/slo-example.pdf){target="_blank"}
* [Example 2](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/examples/africa-example.html){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Identification of points dataset",
                          "Identification of a line dataset",
                          "Identification of a polygon dataset",
                          "Spatial extent (points)",
                          "Unit of analysis (points)",
                          "Number of observations (points)",
                          "Data format (points)",
                          "Variable descriptions (points)",
                          "Coordinate reference system (points)",
                          "Spatial extent (lines)",
                          "Unit of analysis (lines)",
                          "Number of observations (lines)",
                          "Data format (lines)",
                          "Variable descriptions (lines)",
                          "Coordinate reference system (lines)",
                          "Spatial extent (polygons)",
                          "Unit of analysis (polygons)",
                          "Number of observations (polygons)",
                          "Data format (polygons)",
                          "Variable descriptions (polygons)",
                          "Coordinate reference system (polygons)",
                          "Total"),
                 `No credit` = c(rep("Missing", 23), ""),
                 `Half credit` = c(rep("Not adequately described", 23), ""),
                 `Full credit` = c(rep("Adequately described", 23), ""),
                 `Possible points` = c(rep(2, 2), rep(1, 21), 25))

kable(rubric)
```
