---
title: "Plotting spatial data with ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can create beautiful maps using ggplot. The focus of this page 
is to offer a minimal example of how you can quickly plot a spatial 
dataset. This will be useful if you have loaded some data and you 
want to quickly confirm that the file contains what you expected it 
to.

# Required packages

This skills page uses the following three packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(here)
```

# Load data

I'll start by reading in some data You can read more about that 
[here](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-open-file.html){target="_blank"}. 

```{r, message=FALSE, warning=FALSE}
SLO_streets <- st_read(here("a1-files",
                            "SLO-streets.geojson"))

bus_stops <- st_read(here("a1-files",
                          "bus-stops.csv"),
                     options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  st_set_crs("WGS84")
```

# Plot data

You can use the `ggplot()` for a wide variety of data 
visualizations, including scatter plots, bar charts, histograms,
and (what we'll use it for here) maps.

The first line in the code chunk below sets up the plot and 
specifies the dataset the plot will represent. `geom_sf()` indicates
that we want to plot the spatial data as a map.

```{r}
ggplot(SLO_streets) +
  geom_sf()
```

If you want to plot another layer on top of that (perhaps just to
confirm that it has about the same spatial extent), you can call
`geom_sf()` again, but this time you'll need to specify the data
layer within that function. You might also want to change the color
of the second layer to make it easier to see.

```{r}
ggplot(SLO_streets) +
  geom_sf() +
  geom_sf(data = bus_stops,
          color = "red")
```
