---
title: "Spatial filtering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After you've loaded a dataset in R, you may only want to include
certain features in your further analysis, based on feature locations.

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tigris)
library(tidyverse)
library(sf)
library(here)
```

# Loading sample data

In this example, I'm going to start with Rhode Island landmark data 
downloaded from the United States Census Bureau using the `tigris` 
package (see more on getting data from the `tigris` package 
[here](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-tigris.html){target="_blank"}.)

```{r, message = FALSE, results='hide'}
ri_landmark_pts <- landmarks(state = "Rhode Island", type = "point")
```

This gives me all the landmarks (as identified by the Census Bureau) in 
Rhode Island. But I really only want the landmarks within the city
of Providence. I can get the municipal boundary from the city's open
data portal (see more on how to do that here). I'll need to convert the 
coordinate reference system to match the one in my landmarks layer (more
on that here).

```{r, results='hide'}
providence <- st_read("https://data.providenceri.gov/api/geospatial/b6kp-yncs?method=export&format=GeoJSON") %>%
  st_transform(st_crs(ri_landmark_pts))
```

# Filtering data

I can use the following code to create a new dataset called 
`pro_landmark_pts`. I start with `ri_landmark_pts`, then use the 
`st_filter()` function to indicate that I only want to keep the 
points that are within the Providence boundary.

```{r}
pro_landmark_pts <- ri_landmark_pts %>%
  st_filter(providence)
```
