---
title: "Getting Street Data from OpenStreetMap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

OpenStreetMap can be a good sources of street data for many areas of the world.
`osmextract` is one of a few different R packages that make it easy to download
OpenStreetMap data directly into R. YOu can find more details on the package 
and how to use it [here](https://cran.r-project.org/web/packages/osmextract/vignettes/osmextract.html){target="_blank"}.

# Required packages

This skills page uses the following three packages:

```{r, message=FALSE, warning=FALSE}
library(osmextract)
library(tidyverse)
library(sf)
library(here)
```

# Search for a location

All of the functions in the `osmextract` package begin with the prefix `oe_`.We'll use `oe_match()` to find a url for the San Luis Obispo street network. `oe_match()` will save a download url to the variable called `SLO_file` and print a message with the name of the provider.

```{r}
SLO_file <- oe_match("San Luis Obispo")
```

# Download data

Now, we can use `oe_read()` to both download the data to the networks folder and
to read in a layer of lines representing the street network. The function will 
read all line features within the area (including rivers, for example), so we'll
filter to only include lines that are tagged as "highways" (all roads and even 
pedestrian paths are tagged as highways in OpenStreetMap).

```{r}
SLO_streets <- oe_read(SLO_file$url, 
                   provider = "openstreetmap_fr", 
                   layer = "lines", 
                   quiet = TRUE) %>%
  filter(!is.na(highway)) 
```

