---
title: "Saving spatial data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Once you've loaded and/or edited data in R, you might want to 
save it to a file that you can open later in R or another 
software.

# Required packages

This skills page uses the following three packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(here)
```

# Load data from a  file

I'll start by reading in some data. 

```{r, message=FALSE, warning=FALSE}
SLO_streets <- st_read(here("a1-files",
                            "SLO-streets.geojson"))
```

# Save data to a file

The `sf` package has a function 
called `st_write()` that can save your spatial data into most common spatial 
data formats (all functions in the sf package begin with an `st_` prefix).

Here is how I would save my streets as a shapefile I can open in ArcGIS. 
Remember that ESRI shapefiles are collections of files, and it's a
good practice to keep this set of files in its own little subfolder.
The code below will write a set of four files into a folder called 
"SLO-streets" (you will need to have already created a folder with
that name in your working directory).

```{r, eval=FALSE}
st_write(SLO_streets, here("SLO-streets", 
                           "SLO-streets.shp"))
```

You could use the same function to save it as a geoJSON file. In this case, you 
don't need to make a subfolder for it first because a geoJSON is a single file.

```{r, eval=FALSE}
st_write(SLO_streets, here("SLO-streets.geojson"))
```
