---
title: "Opening spatial data files in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can download spatial data files from a variety of sources, to 
analyze and display your data in R, you'll need to open these
files in R

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(here)
```

# Opening a file

The function `st_read()` will open spatial data from almost any
common data format. For shapefiles, you can use the name of the
folder containing the set of files, or the name of the file with the
*.shp extension (as long as the other files are in the same directory
and have the same root name).

```{r}
boston_bikes <- st_read(here("a1-files",
                             "Existing_Bike_Network"))
```

For other file formats that just use a single file, use the filename.

```{r}
SLO_streets <- st_read(here("a1-files",
                            "SLO-streets.geojson"))
```

