---
title: "Opening spatial data files in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can download spatial data files from a variety of sources, to 
analyze and display your data in R, you'll need to open these
files in R

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(here)
```

# Opening a spatial data file

The function `st_read()` will open spatial data from almost any
common data format. For shapefiles, you can use the name of the
folder containing the set of files, or the name of the file with the
*.shp extension (as long as the other files are in the same directory
and have the same root name).

```{r, results='hide'}
boston_bikes <- st_read(here("a1-files",
                             "Existing_Bike_Network"))
```

For other file formats that just use a single file, use the filename.

```{r, results='hide'}
SLO_streets <- st_read(here("a1-files",
                            "SLO-streets.geojson"))
```

# Opening a CSV file

A csv file is not exactly a spatial data file. It is basically
a text file with tabular data, where rows are in separate
lines and columns are separated by columns. Sometimes point
data will be stored in a csv file, with a column for an 
x-coordinate (longitude) and a column for a y-coordinate 
(latitude). 

The `st_read()` function can also read point data 
from a csv file, but you'll need to set some options indicating
which columns are actually coordinates. You'll also need to
specify a coordinate reference system (other spatial data formats
will include this information, but a csv file will not). If you 
know the coordinates are latitude/longitude values, WGS 84 is a pretty safe
CRS to guess, if you don't have information indicating otherwise.

```{r, results='hide'}
bus_stops <- st_read(here("a1-files",
                          "bus-stops.csv"),
                     options=c("X_POSSIBLE_NAMES=stop_lat",
                               "Y_POSSIBLE_NAMES=stop_lon")) %>%
  st_set_crs("WGS84")
```

