---
title: "Getting national and subnational boundaries from rnaturalearth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`rnaturalearth` is a useful package for loading data from 
[Natural Earth](https://www.naturalearthdata.com/).

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(rnaturalearthdata)
library(rnaturalearth)
library(tidyverse)
library(sf)
library(here)
```

# Loading country boundaries

The `ne_countries()` function will load national boundaries.

Here is how you would load a set of 177 countries available from rnaturalearth.

```{r}
world <- ne_countries(returnclass = "sf")
```

Here is how you would load the boundaries of South Africa, Namibia, Lesotho, and Botswana.

```{r}
s_africa <- ne_countries(country = c("South Africa", 
                                  "Namibia",
                                  "Botswana",
                                  "Lesotho"),
                      returnclass = "sf")
```

Here is how you would load the boundaries of all the countries in South America.

```{r}
s_africa <- ne_countries(continent = "South America",
                      returnclass = "sf")
```

# Loading state/province boundaries

The `ne_states()` function returns state/provincial boundaries (administrative level 1).

Here's how you would load the boundaries of Canadian provinces. 

Note: you may need to install rnaturalearthhires for this to work. You can install it from GitHub
by typing `devtools::install_git("https://github.com/ropensci/rnaturalearthhires")` into your 
console. You'll only need to ever do this once.

```{r}
canada_prov <- ne_states(country = "Canada",
                         returnclass = "sf")
```





