---
title: "Changing a dataset's coordinate reference system"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Spatial data can be stored using one of many coordinate reference 
systems (CRS). If you need two layers of data to interact with one
another, they need to be in the same CRS. Moreover, some packages and 
functions in R require your data to be in a particular CRS (often WGS 
84). So it's helpful to know how to convert your data from one CRS to 
another.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tigris)
library(sf)
library(here)
```

# Common geographic coordinate reference systems

WGS 84, NAD 83, and NAD 27 are common coordinate reference systems for 
data stored as latitude and longitude coordinates (geographic coordinate
systems). You can use the function `st_transform()` to convert any spatial
data set into any of these.

In this code chunk, I'll read in a spatial dataset from a file and then use 
`st_transform()` to convert it to NAD 83. To convert it to WGS 84, use 
`"WGS84"` instead of `"NAD83"`. To convert it to NAD 27, use `"NAD27"`.

```{r results='hide'}
streets <- st_read(here("a1-files",
                        "SLO-streets.geojson")) %>%
  st_transform("NAD83")
```

# Matching coordinate reference systems

In some cases, you may not really care what CRS your data is in, but you 
do need it to be consistent across layers.

In this example, I'll load a set of census tracts using the tigris package
(more on that here) and set the CRS to match the CRS in the streets layer.

```{r, message=FALSE, results='hide'}
tracts <- tracts(state = "CA",
                 county = "San Luis Obispo") %>%
  st_transform(st_crs(streets))
```

