---
title: "Loading spatial data from tigris"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `tigris` package is a convenient way to load TIGER data from 
the United States Census Bureau into R. The author of the `tigris`
package, Kyle Walker, has written a free ebook about working with 
census data in R, and you can read more about the `tigris` package
in chapter 5 of that book, [here](https://walker-data.com/census-r/census-geographic-data-and-applications-in-r.html).

Census boundaries, such as national, state, county, or census tract
boundaries can be loaded from `tigris`, but if you load them through
the `tidycensus` package, you can get a wide variety of population
and housing variables at the same time, so I recommend you use 
`tidycensus` for census boundaries. `tigris` is especially useful 
for loading data on roads, water features, and landmarks.

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tigris)
library(tidyverse)
library(sf)
library(here)
```

# Getting and installing a census API key

Before you can use the `tidycensus` package, you will need to 
register for a  Census API key
[here](https://api.census.gov/data/key_signup.html){target="_blank"}. 
It can take one or two business days to get a key, *so do 
this well before the assignment is due*.

Your API key will be a string of letters and numbers. If your API key is 
"1234567abcdEFG" (it isn't), you can install your API key by typing 
`census_api_key("1234567abcdEFG", install = TRUE)` into your console. 
You should only ever need to do this once.

# Loading data

The census collects data on landmarks such as parks, schools, 
airports, and cemeteries. Some landmarks are stored as points, and 
others are stored as areas. You can use the `landmarks()` function 
to download either point landmarks or area landmarks. Landmarks are 
available for states (i.e. you must specify which state you want
and you can't specify a lower geography level like a county).

Here is how you would download all the point landmarks in Rhode 
Island:

```{r, message = FALSE, results='hide'}
ri_landmark_pts <- landmarks(state = "Rhode Island", type = "point")
```

Here is how I would download all the area landmarks in Rhode Island in
Connecticut:

```{r, message = FALSE, results='hide'}
ct_landmark_areas <- landmarks(state = "Connecticut", type = "area")
```

`area_water()` represents areas of water such as lakes or areas 
covered by rivers and streams as polygons. You must specify a state and
a county.

```{r, message = FALSE, results='hide'}
boston_lakes <- area_water(state = "Massachusetts", county = "Suffolk")
```

`linear_water()` represents rivers, streams, and canals as lines. You
must speficy a state and a county.

```{r, message = FALSE, results='hide'}
westchester_rivers <- linear_water(state = "New York", 
                                   county = "Westchester")
```

`primary_roads()` represents major roads (e.g. freeways) as lines. 
You can't specify a specific geography; this function will give you all
the primary roads in the United States.

```{r, message = FALSE, results='hide'}
usa_roads <- primary_roads()
```

`primary_secondary_roads()` represents roads like freeways and 
arterials (primary and secondary roads) as lines. You must specify a 
state.

```{r, message = FALSE, results='hide'}
kentucky_roads <- primary_secondary_roads(state = "Kentucky")
```

`roads()` represents all roads, including local neighborhood streets, as lines. You mush specify a state and county.

```{r, message = FALSE, results='hide'}
SLC_roads <- roads(state = "Utah", county = "Salt Lake")
```

`rails()` represents all rail lines as as lines. You can't specify a specific geography; this function will give you all the 
rail lines in the United States.

```{r, message = FALSE, results='hide'}
USA_rail <- rails()
```

# Differentiating among types of features

TIGER features (whether you download them using `tigris` or via the 
web interface) will include a variable called MTFCC with a five-digit 
code (a letter followed by four numbers). These codes classify features
by type. For example, the boston_lakes dataset loaded above using the
`area_water()` function has four different values for this variable: 
H2030, H2040, H2051, and H3010. You can find definitions of these
codes at the census website, 
[here](https://www.census.gov/library/reference/code-lists/mt-feature-class-codes.html){target="_blank"}, where you'll see that:

* H2030 indicates a lake/pond
* H2040 indicates a resevoir
* H2051 indicates a bay/estuary/gulf/sound
* H3010 indicates a stream/river