---
title: "Aspatial filtering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After you've loaded a dataset in R, you may only want to include
certain features in your further analysis, based on the values
of a variable.

# Required packages

This skills page uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tigris)
library(tidyverse)
library(sf)
library(here)
```

# Loading sample data

In this example, I'm going to start with Rhode Island landmark data 
downloaded from the United States Census Bureau using the `tigris` 
package (see more on getting data from the `tigris` package 
[here](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-tigris.html){target="_blank"}.)

```{r, message = FALSE, results='hide'}
ri_landmark_pts <- landmarks(state = "Rhode Island", type = "point")
```

TIGER features (whether you download them using `tigris` or via the 
web interface) will include a variable called MTFCC with a five-digit 
code (a letter followed by four numbers). These codes classify features
by type. The landmark dataset loaded above using the
has 16 different values for this variable, which I can list using the 
following code.

I start with the full dataset, then the `st_drop_geometry()` drops the 
geometry column, and the `select()` function drops all the other columns
except MTFCC. `distinct()` filters to only include the unique rows of
this new one-variable dataset. Note that I am not editing the dataset;
I'm creating a new dataframe and displaying its contents without saving 
it anywhere.

```{r}
ri_landmark_pts %>%
  st_drop_geometry() %>%
  select(MTFCC) %>%
  distinct()
```

You can find definitions of the MTFCC codes at the census website, 
[here](https://www.census.gov/library/reference/code-lists/mt-feature-class-codes.html){target="_blank"}, 
where you'll see that K2582 indicates a cemetery. 

# Filtering data

I can use the following code to create a new dataset called 
`ri_cemeteries`. I start with `ri_landmark_pts`, then use the 
`filter()` function to indicate that I only want to keep the 
rows where the value of MTFCC is K2582 (the code for a cemetery).

```{r}
ri_cemeteries <- ri_landmark_pts %>%
  filter(MTFCC == "K2582")
```
