---
title: "Viewing a dataset as a table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The easiest way to figure out what the rows and columns in your 
dataset represent will often be to view the rows and columns in 
a table.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tigris)
library(here)
```

# Loading example data

I'll start by loading a set of census tracts using the `tigris`
package.

```{r, message=FALSE, results='hide'}
tracts <- tracts(state = "Illinois",
                 county = "Cook") 
```

# Viewing your data

Once you've loaded the data, you'll see its name on your environment
tab in RStudio, along withe number of rows (observations or obs.) and the
number of columns (variables)

```{r env-pane, echo = FALSE, fig.asp=1.0, out.width='100%'}
here("a1-files",
     "images",
     "environment-pane.png") %>%
knitr::include_graphics()
```

If you click the blue icon next to the name of the data set, you'll see
a list of the variables, and values for the first few rows.

```{r env-pane2, echo = FALSE, fig.asp=1.0, out.width='100%'}
here("a1-files",
     "images",
     "environment-pane2.png") %>%
knitr::include_graphics()
```

If you click on the name of the data set, RStudio will open a tab showing the
data in a table.

```{r view-data, echo = FALSE, fig.asp=1.0, out.width='100%'}
here("a1-files",
     "images",
     "view-data.png") %>%
knitr::include_graphics()
```

You can also open that tabular view by typing `View(tracts)` into your
console (note that the V is capitalized).
