---
title: "Assignment 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Generate a map to communicate the distances between features
using buffers and/or by symbolizing features based on 
their distance from other features. Your map layout should 
include a description of the analysis question your
map is meant to address and a clear description of what each 
layer represents (this might be in the map legend). Do not 
include a story about how you made the map (e.g. "I generated 
a set of buffers") or the technology you used. Focus on the 
what the map shows and what conclusions you can draw from it.

Include the sources of all the data shown
in your map (including any data that isn't shown but was 
used to derive the layers you've shown), the name of a
classmate who was helpful to you, and how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this 
assignment, but all of them are relevant and you will 
probably use some of them.

#### R

* [Calculating distances between features across layers](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/skill-distance-two-layers.html){target="_blank"}
* [Calculating distances between features within layers](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/skill-distance-one-layer.html){target="_blank"}
* [Generating buffers](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/skill-buffer.html){target="_blank"}
* [Generating centroids](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/skill-centroids.html){target="_blank"}


#### ArcGIS Pro

* [Generating buffers](https://harvard.zoom.us/rec/share/TOxOlJud82yY5PiUdBkOF85xxFJ4crNmWdX7XYsZW627o9yZLHAaOSOdZYXluTTZ.CAPFos-fbWXcBS9R?){target="_blank"}
* [Generating centroids](https://harvard.zoom.us/rec/share/TOxOlJud82yY5PiUdBkOF85xxFJ4crNmWdX7XYsZW627o9yZLHAaOSOdZYXluTTZ.CAPFos-fbWXcBS9R?startTime=1663593420000){target="_blank"}
* Measuring distance (coming soon)
 
# Other resources

* [Buffers in ArcGIS Pro](https://pro.arcgis.com/en/pro-app/latest/tool-reference/analysis/buffer.htm){target="_blank"}
* [Centroids in ArcGIS Pro](https://pro.arcgis.com/en/pro-app/2.9/tool-reference/data-management/feature-to-point.htm){target="_blank"}
* [Centroids in ArcGIS Online](https://doc.arcgis.com/en/arcgis-online/analyze/find-centroids.htm){target="_blank"}

# Examples

The example below fully meets the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a4-files/Examples/slo-example.pdf){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Illustration of distance",
                          "Description of analysis question",
                          "Description of conclusion",
                          "Source attribution",
                          "Total"),
                 `No credit` = c(rep("Missing", 6), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "Unintelligible",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described",
                                   ""),
                 `Full credit` = c("Adequately described",
                                   "Adequately described",
                                   "Distance is legible",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 10, 5, 5, 1, 25))

kable(rubric)
```
