---
title: "Assignment 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Generate the following new data layers:

* A polygon layer representing the area within a constant 
(Euclidean) distance from one or more features in a point 
or line layer.
* A polygon layer representing the area within a constant
travel time (by a non-motorized mode) from one or more 
features in a point layer
* A polygon layer representing the area within a constant
travel time (by a motorized mode) from one or more
features in a point layer
* A layer of points randomly distributed within the 
polygons in a polygon layer
* A layer of points representing the centroids of the
polygons in a polygon layer
* A raster layer representing the distance to the nearest
point in a point dataset from any location in your
study area

Submit two layouts (either as separate files or as two
pages of the same file). 

The first layout should address a specific question 
with a map that includes one or more of the layers 
described above The layout should include text that 
describes the question your map is meant to answer. The
second layout should illustrate any layers 
described above that aren't included in your first layout 
(just to demonstrate that you know how to generate them).

On your submission, list the name(s) of at least one 
student who was especially helpful you you in completing 
this assignment, and describe how they helped. You should also clearly indicate the sources of all your data.

# Relevant Skills

You are unlikely to use all of these skills in this 
assignment, but all of them are relevant and you will 
probably use some of them.

#### R

* Generating buffers
* Generating isochrones
* Generating centroids
* Spatial sampling
* Generating a grid
* Generating a raster from a grid
* Generating a matrix of Euclidean distances between
two point layers
* Generating a matrix of travel times between two 
point layers

#### ArcGIS

* Generating buffers
* Generating isochrones
* Generating centroids
* Spatial sampling
* Generating a grid
* Generating a raster from a grid
* Generating a matrix of Euclidean distances between
two point layers
* Generating a matrix of travel times between two 
point layers

# Other resources

* [r5r documentation on building travel time matrices](https://ipeagit.github.io/r5r/articles/travel_time_matrix.html){target="_blank"}

# Examples

Each of the examples below fully meets the requirements for this assignment.

* [Example 1]
* [Example 2]

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Inclusion of a Euclidean buffer layer",
                          "Inclusion of an isochrone for a motorized mode",
                          "Inclusion of an isochrone for a non-motorized mode",
                          "Inclusion of a layer of random points",
                          "Inclusion of a layer of centroids",
                          "Inclusion of a distance raster",
                          "Source attribution",
                          "Description of research question",
                          "Total"),
                 `No credit` = c(rep("Missing", 10), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   rep("",6),
                                   "Not adequately described",
                                   "Not adequately described", 
                                   ""),
                 `Full credit` = c("Adequately described",
                                   "Adequately described",
                                   rep("Included",6),
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(3, 3, 2, 2, 2, 2, 2, 2, 3, 4, 25))

kable(rubric)
```
