---
title: "Generating a buffer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A buffer is a polygon layer that indicates the difference with
a specified distance of a set of features.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(here)
library(leaflet)
```

# Load some sample data

```{r, include=FALSE}

library(tigris)

co_trails <- here("a1-files",
                  "County_Trails.geojson") %>%
  st_read() %>%
  dplyr::select(TrailName, Length, LengthUnits)

city_trails <- here("a2-files",
                    "Examples",
                    "Trail.geojson") %>%
  st_read() %>%
  rename(TrailName = Trail_Name,
         Length = SLength) %>%
  mutate(LengthUnits = "MILES") %>%
  dplyr::select(TrailName, Length, LengthUnits)

trails <- rbind(city_trails, co_trails) 

rta_stops <- here("a1-files",
                  "bus-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "San Luis Obispo RTA") 

slo_stops <- here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "SLO Transit") %>%
  dplyr::select(colnames(rta_stops)) 

stops <- rbind(rta_stops, slo_stops) %>%
  st_set_crs("WGS84") %>%
  mutate(id = 
           as.character(seq(1,
                            length(rta_stops$stop_id) + 
                              length(slo_stops$stop_id),
                            by=1))) 

city <- places(state = "CA") %>%
  filter(NAME == "San Luis Obispo") %>%
  st_transform("WGS84")

```

I'll demonstrate buffers from points, lines, and polygons. 

I have a layer called `stops` that shows the locations of
bus stops in San Luis Obispo County, a layer called `trails`
that shows all city- and county- managed trails, and a layer 
of tracts that shows the census tract boundaries. Here's
a quick map of all three layers.

```{r}
leaflet() %>%
  addProviderTiles(provider = providers$Stamen.TerrainBackground) %>%
  setView(lng = -120.6592, lat = 35.28190, zoom = 12) %>%
  addCircleMarkers(data = stops,
                   radius = 2,
                   stroke = FALSE,
                   fillColor = "darkorange",
                   fillOpacity = 1) %>%
   addPolylines(data = trails,
                color = "brown",
                weight = 1.5,
                opacity = 1) %>%
   addPolygons(data = city,
               fill = NA,
               color = "black",
               opacity = 1,
               weight = 2)
```  

# Transform the data

To generate accurate buffers, my data need to be stored in a 
projected coordinate system (with coordinates in units of distance).

I'll use the California State Plane, zone V (https://spatialreference.org/ref/esri/102645/),
which stores coordinates in units of feet.

```{r}
# Define CA-state-plane
CA_st_plane <- "+proj=lcc +lat_1=34.03333333333333 +lat_2=35.46666666666667 +lat_0=33.5 +lon_0=-118 +x_0=2000000 +y_0=500000.0000000002 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

trails_st_plane <- trails %>%
  st_transform(CA_st_plane)

stops_st_plane <- stops %>%
  st_transform(CA_st_plane)

city_st_plane <- city %>%
  st_transform(CA_st_plane)
```

# Point Buffers

The `st_buffer()` function will generate a buffer around a each 
feature in your dataset. For a layer of points, this will be a pile 
of circles, each centered on a stop. The `dist` parameter is the 
radius of each circle, in the same units as the coordinate system.

The Califonia State Plane system I'm using is in feet, so I need
to specify the distance in feet. I'll use 10560 feet (two miles).

```{r}
stop_buffers <- stops_st_plane %>%
  st_buffer(dist = 10560) 

ggplot() +
  geom_sf(data = stop_buffers) +
  theme_void()
```

I might want a polygon representing the area within a mile of _any_ 
transit stop, rather than a separate circle for each stop. Combining 
overlapping features into a single feature is called a spatial union,
and I can do this using the `st_union()` function.

```{r}
stop_buffer_combined <- stops_st_plane %>%
  st_buffer(dist = 10560) %>%
  st_union()

ggplot() +
  geom_sf(data = stop_buffer_combined) +
  theme_void()
```

# Line buffers

The `st_buffer` function works the same way for lines. Here's
the area within a half mile of a trail.

```{r}
trail_buffer <- trails_st_plane %>%
  st_buffer(dist = 2640) %>%
  st_union()

ggplot() +
  geom_sf(data = trail_buffer) +
  geom_sf(data = trails) +
  theme_void()
```

# Polygon buffers

And you can also create a buffer around a polygon. Here's
the area within a half mile of the San Luis Obispo city limits.

```{r}
city_buffer <- city_st_plane %>%
  st_buffer(dist = 2640)

ggplot() +
  geom_sf(data = city_buffer,
          fill = "pink") +
  geom_sf(data = city_st_plane,
          fill = "yellow") +
  theme_void()
```