---
title: "Generating random points within a polygon"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are some cases where you might be interested in the distance
from a points that are evenly distributed throughout a polygon. Here's
how you would generate a grid of points.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(tidycensus)
```

# Load some sample data

```{r, include = FALSE}
# Define CA-state-plane
CA_st_plane <- "+proj=lcc +lat_1=34.03333333333333 +lat_2=35.46666666666667 +lat_0=33.5 +lon_0=-118 +x_0=2000000 +y_0=500000.0000000002 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"


tracts <- get_decennial(year = 2020,
                        geography = "tract", 
                        variables = c(Population = "P2_001N"), 
                        output = "wide", 
                        state = "CA",
                        county = "San Luis Obispo",
                        geometry = TRUE) %>%
   st_transform(CA_st_plane)


```

I'll start with this map of census tracts in San Luis Obispo County. The data are in the 
California State Plane coordinate system.

```{r}
ggplot(tracts) +
  geom_sf()
```  

# Generate a grid of points

The `st_make_grid()` function will generate a grid that covers the bounding box of a
specified layer.

```{r}
grid_squares <- st_make_grid(tracts)

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_squares,
          fill = NA)
```

By default, the grid sizes are selected so to make a 10x10 grid. I can also specific the
number of grid cells I want in each direction. Here is a 15x12 grid.

```{r}
grid_squares <- st_make_grid(tracts, n = c(15, 12))

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_squares,
          fill = NA)
```

It's more likely that you'll care more about the cell size than about how many cells 
are in the bounding box. You can also specify the cell height and width instead of the 
number of cells. Here is a grid of cells that are 5 miles by 5 miles (there are 5280 
feet in a mile).

```{r}
grid_squares <- st_make_grid(tracts, cellsize = c(5 * 5280, 5 * 5280))

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_squares,
          fill = NA)
```

You can also make a hexagonal grid by setting `square = FALSE`.

```{r}
grid_hex <- st_make_grid(tracts, cellsize = c(5 * 5280, 5 * 5280), square = FALSE)

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_hex,
          fill = NA)
```

If you want points instead of polygons, you can set `what = corners` or `what = centers`.

Here is a set of points at the centers of a 5-mile by 5-mile hex grid. 

```{r}
grid_hex_pts <- st_make_grid(tracts, 
                             cellsize = c(5 * 5280, 5 * 5280), 
                             square = FALSE,
                             what = "centers")

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_hex_pts,
          fill = NA)
```

Here's a hex grid with points at the centers of square grid cells.

```{r}
grid_hex_pts <- st_make_grid(tracts, 
                             cellsize = c(5 * 5280, 5 * 5280), 
                             what = "corners")

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_hex_pts,
          fill = NA)
```

Here is a set of points at the corners of a hex grid. 

```{r}
grid_hex_pts <- st_make_grid(tracts, 
                             cellsize = c(5 * 5280, 5 * 5280), 
                             square = FALSE,
                             what = "corners")

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_hex_pts,
          fill = NA)
```

In this case, the layer of points was automatically clipped to the polygon boundary, where
all the other grids filled the whole bounding box. I do not have an explanation for that.

You can easily clip any grid to the boundaries of a polygon though.

```{r}
grid_hex_pts <- st_make_grid(tracts, 
                             cellsize = c(5 * 5280, 5 * 5280), 
                             what = "corners") %>%
  st_as_sf() %>%
  st_filter(tracts)

ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = grid_hex_pts,
          fill = NA)
```


