---
title: "Calculating distances within a layer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You for a set of features, you can calculate a variable
that indicates the difference to the closest feature
within the dataset.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(here)
library(leaflet)
library(knitr)
```

# Load some sample data

```{r, include=FALSE}
rta_stops <- here("a1-files",
                  "bus-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "San Luis Obispo RTA") 

slo_stops <- here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "SLO Transit") %>%
  dplyr::select(colnames(rta_stops)) 

stops <- rbind(rta_stops, slo_stops) %>%
  st_set_crs("WGS84") %>%
  mutate(id = 
           as.character(seq(1,
                            length(rta_stops$stop_id) + 
                              length(slo_stops$stop_id),
                            by=1))) %>%
  select(stop_id, stop_name)

```

I'll demonstrate this with a layer of transit stop
locations (`stops`) in San Luis Obispo County, California. 

```{r}
leaflet() %>%
  addProviderTiles(provider = providers$Stamen.TerrainBackground) %>%
  addCircleMarkers(data = stops,
                   radius = 2,
                   stroke = FALSE,
                   fillColor = "darkorange",
                   fillOpacity = 1) 
```  

# Transform the data

If I want to calculate distances, my data need to be stored in a 
projected coordinate system (with coordinates in units of distance).

I'll use the California State Plane, zone V (https://spatialreference.org/ref/esri/102645/),
which stores coordinates in units of feet.

```{r}
# Define CA-state-plane
CA_st_plane <- "+proj=lcc +lat_1=34.03333333333333 +lat_2=35.46666666666667 +lat_0=33.5 +lon_0=-118 +x_0=2000000 +y_0=500000.0000000002 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

stops_st_plane <- stops %>%
  st_transform(CA_st_plane)
```

# Calculating the distance to the nearest feature within a layer

The `st_nearest_feature()` function will return the index 
(the row number) of the nearest point in the layer, for each
point in the layer.

```{r}
nearest_stop <- st_nearest_feature(stops_st_plane)
```

Now you can use `st_distance()` to calculate the distance between each stop and 
the nearest stop. The first argument in the `st_distance()` function below
is the layer of points. The second argument is the layer of points again,
reordered according to the indices in the `nearest_stop` vector I created
above. `by_element = TRUE` means I want one distance for each point in the layer
(rather than the distance from each point to every other point).

```{r}
stops_st_plane <- stops_st_plane %>%
  mutate(distance_to_nearest = 
           st_distance(stops_st_plane, 
                       stops_st_plane[nearest_stop,],
                       by_element = TRUE))
```

Here are the first few rows of the resulting data frame, which now
includes a column with the distance to the nearest other stop.

```{r}
head(stops_st_plane) %>%
  kable()
```

You'll see that the distance is very helpfully stored in a data type that
labels the units. This is cool, but it actually can make subsequent
steps you might take a little trickier. So let's change these to regular
numbers.

```{r}
stops_st_plane <- stops_st_plane %>%
  mutate(distance_to_nearest = as.numeric(distance_to_nearest))

head(stops_st_plane) %>%
  kable()
```

# Map showing variation in distances

Now I can show the results on a map. You can see that the more 
isoloated stops are shown in a lighter color.

```{r}
ggplot(stops_st_plane) +
  geom_sf(aes(color = distance_to_nearest),
          size = 1.5) 
```
