---
title: "Calculating distances across layers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You for a set of features, you can calculate a variable
that indicates the difference to the closest feature
in another layer

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(here)
library(leaflet)
library(knitr)
```

# Load some sample data

```{r, include=FALSE}
rta_stops <- here("a1-files",
                  "bus-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "San Luis Obispo RTA") 

slo_stops <- here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                    "Y_POSSIBLE_NAMES=stop_lat")) %>%
  mutate(`Transit Provider` = "SLO Transit") %>%
  dplyr::select(colnames(rta_stops)) 

stops <- rbind(rta_stops, slo_stops) %>%
  st_set_crs("WGS84") %>%
  mutate(id = 
           as.character(seq(1,
                            length(rta_stops$stop_id) + 
                              length(slo_stops$stop_id),
                            by=1))) %>%
  select(stop_id, stop_name)

co_trails <- here("a1-files",
               "County_Trails.geojson") %>%
  st_read() %>%
  select(TrailName, Length, LengthUnits)

city_trails <- here("a2-files",
                    "Examples",
               "Trail.geojson") %>%
  st_read() %>%
  rename(TrailName = Trail_Name,
         Length = SLength) %>%
  mutate(LengthUnits = "MILES") %>%
  select(TrailName, Length, LengthUnits)

trails <- rbind(city_trails, co_trails) 

```

I'll demonstrate this with a layer of transit stop
locations (`stops`) and trails (`trails`) in San Luis 
Obispo County, California. 

```{r}
leaflet() %>%
  addProviderTiles(provider = providers$Stamen.TerrainBackground) %>%
  addCircleMarkers(data = stops,
                   radius = 2,
                   stroke = FALSE,
                   fillColor = "darkorange",
                   fillOpacity = 1) %>%
  addPolylines(data = trails,
               weight = 1,
               color = "black",
               opacity = 1)
```  

# Transform the data

If I want to calculate distances, my data need to be stored in a 
projected coordinate system (with coordinates in units of distance).

I'll use the California State Plane, zone V (https://spatialreference.org/ref/esri/102645/),
which stores coordinates in units of feet.

```{r}
# Define CA-state-plane
CA_st_plane <- "+proj=lcc +lat_1=34.03333333333333 +lat_2=35.46666666666667 +lat_0=33.5 +lon_0=-118 +x_0=2000000 +y_0=500000.0000000002 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

stops_st_plane <- stops %>%
  st_transform(CA_st_plane)

trails_st_plane <- trails %>%
  st_transform(CA_st_plane)
```

# Calculating the distance to the nearest feature within a layer

The `st_nearest_feature()` function will return the index 
(the row number) of the nearest point in the second layer, for each
point in the first layer. This example gives the index of the 
nearest stop to each trail.

```{r}
nearest_stop <- st_nearest_feature(trails_st_plane,
                                   stops_st_plane)
```

Now you can use `st_distance()` to calculate the distance between each trail and 
the nearest stop. `by_element = TRUE` means I want one distance for each trail
(rather than the distance from each each trail to each stop).

```{r}
trails_st_plane <- trails_st_plane %>%
  mutate(distance_to_transit = 
           st_distance(trails_st_plane, 
                       stops_st_plane[nearest_stop,],
                       by_element = TRUE))
```

Here are the first few rows of the resulting data frame, which now
includes a column with the distance to the nearest other stop.

```{r}
head(trails_st_plane) %>%
  kable()
```

You'll see that the distance is very helpfully stored in a data type that
labels the units. This is cool, but it actually can make subsequent
steps you might take a little trickier. So let's change these to regular
numbers.

```{r}
trails_st_plane <- trails_st_plane %>%
  mutate(distance_to_transit = as.numeric(distance_to_transit))

head(trails_st_plane) %>%
  kable()
```

# Map showing variation in distances

Now I can show the results on a map. You can see that the trails that
are farther from transit are shown in lighter color.

```{r}
ggplot() +
    geom_sf(data = stops_st_plane,
          size = 1,
          color = "gray") +
  geom_sf(data = trails_st_plane,
          aes(color = distance_to_transit)) +
  scale_color_viridis_c() +
  theme_void()
```
