---
title: "Generating centroids"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sometimes I'll be more interested in the distance to the middle
of a something than I am in the distance to the closest part of it.
In those cases, it's useful to generate a centroid.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
```

# Load some sample data

```{r, include=FALSE}
library(tidycensus)

tracts <- get_acs(geography = "tract", 
                  variables = "B06011_001", 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) 

```

Here are the census tracts in San Luis Obispo, California

```{r}
ggplot(tracts) +
  geom_sf()
```  

# Generate centroids

I can use `st_centroid()` to generate a layer of points representing the
centroid of each tract. `st_centroid()` always generates a warning about 
an assumption of constant attributes. This is fine.

```{r}
tract_centroids <- st_centroid(tracts) 
```

# Map

Here is a quick map showing the centroid of each tract.

```{r}
ggplot() +
  geom_sf(data = tracts) +
  geom_sf(data = tract_centroids) 
```
