---
title: "Calculating new variables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Any data analysis will likely require you to calculate new variables.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(knitr)
```

# Load some sample data

R has several datasets built in for practice and testing. I'll start by
loading on of those with statistics on violent crime rates by state in 1973. 
It includes four variables:

* *Murder*: The number of murder arrests per 100,000 residents
* *Assault*: The number of assault arrests per 100,000 residents
* *UrbanPop*: The percent of the state's population in urban areas
* *Rape*: The number of rape arrests per 100,000 residents

```{r}
# Load the dataset
data("USArrests")

# Display the first few rows in a table
head(USArrests) %>%
  kable()
```

# The mutate function

The total number of crimes in a state would be the sum of the murder, assault, and rape
rates, times 100,000. I can use the `mutate()` function to calculate this value and save 
it in a new variable called `crime_count`.

```{r}
# Calculate a new variable
USArrests <- USArrests %>%
  mutate(crime_count = (Murder + Assault + Rape) * 100000)

# Display the first few rows in a table (format the large numbers)
head(USArrests) %>%
  kable(format.args = list(big.mark = ","))
```


