---
title: "Boston's Changing Shoreline"
output: 
  tufte::tufte_html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Much of the Boston's land area was constucted from "filled land". 
This map, created in 1895 by Boston City
Surveyor, Charles Carroll Perkins shows the orginal shoreline in gray and 
the changing wharf lines from 1795, 1850, and 1895 as the city expanded 
into the bay.

Without mitigation, the rising sea levels that accompany a changing climate may 
reclaim some of land that was created in the first centuries of Boston's 
development. 

This map is intended for Boston residents who are curious about the history
and possible futures of their city.

```{r load-packages, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(here)
```

```{r load-data, include=FALSE}
hi_tide_36_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation.geojson") %>%
  st_read()

hi_tide_21_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation-21in.geojson") %>%
  st_read()

hi_tide_9_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation-9in.geojson") %>%
  st_read()
```

```{r, echo=FALSE}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  addPolygons(data = hi_tide_36_inch,
              weight = 1,
              group = "High tide with 36-inch sea level rise (2070s)") %>%
  addPolygons(data = hi_tide_21_inch,
              weight = 1,
              group = "High tide with 21-inch sea level rise (2050s)") %>%
    addPolygons(data = hi_tide_9_inch,
              weight = 1,
              group = "High tide with 9-inch sea level rise (2030s)") %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15) %>%
  addLayersControl(baseGroups = 
                     c("High tide with 36-inch sea level rise (2070s)",
                       "High tide with 21-inch sea level rise (2050s)",
                       "High tide with 9-inch sea level rise (2030s)"),
    options = layersControlOptions(collapsed = FALSE)) 
```

#### *Date Sources:*

* Perkins, & Geo. H. Walker & Co. (1895). Plan of Boston proper: showing changes in street and wharf lines, 1795 to 1895 [Map]. Geo. H. Walker, lith. https://digitalcollections.library.harvard.edu/catalog/990094068540203941
* Climate Ready Boston Sea Level Inundation. Analyze Boston. https://data.boston.gov/dataset/climate-ready-boston-sea-level-rise-inundation

#### *Acknowldgments:*

My classmate, Belle Lipton, helped me with this assignment by showing me how to use
Allmaps to generate map tiles for Leaflet.