---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Submit a map with showing the data you described in 
Assignment 1, with each dataset as a separate layer. 

Your map should be attractive and informative (inset maps, north arrows, and 
scale bars are very helpful in making a map informative, but I'm not 
dogmatic about including them if your audience doesn't need them). It should 
clearly indicate the source of each layer in your map. It should
also include a couple of sentences describing the audience for your map and
the argument you are making or the message you are conveying to them with
your map.

On your submission, list the name(s) of at least one student who was 
especially helpful you you in completing this assignment, and describe 
how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this assignment, but all 
of them are relevant and you will probably use several of them. They are
listed alphabetically.

* [Adding a map layout (ArcGIS)](https://harvard.zoom.us/rec/play/NXs6WI6aTbaq1VwV6jJL9wWGBLpweSkK9t9ktZQ1ambeUJD0Kk6stBrrVUxngXFBsdXQJjSrDVHTqtU5.UzDytkbnGdiTaZzO?autoplay=true&startTime=1660838502000){target="_blank"}
* [Add a north arrow, scale bar, and legend (ArcGIS)](https://harvard.zoom.us/rec/play/yA2FG1jRi-shr6JTJGxvFRxiwzhVahGMSR0w6zGUeecEf2BdWbzSJENjDMPwcnfHS7_o7BZdwtLeLQmz.7Nn3qaXIs_65VHpM?autoplay=true&startTime=1660838907000){target="_blank"}
* [Adding a north arrow to a map (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-north-arrow.html){target="_blank"}
* [Adding a scale bar to a map (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-scale-bar.html){target="_blank"}
* [Annotating a ggplot map (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-annotate.html){target="_blank"}
* [Applying a ggplot theme (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-themes.html){target="_blank"}
* [Changing symbology based on a variable (ArcGIS)](https://harvard.zoom.us/rec/play/YqECB4uxFWT9P00fxHZYNvV1ekgt4CiINjE3re55aidcxBbakYoJyJsh2uy-GOhiq4xepkJjcqIq1rb1.vvtSc2XTOJvPRrQx?autoplay=true&startTime=1660836764000){target="_blank"}
* [Changing symbology based on a variable (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-variable-symbology.html){target="_blank"}
* [Changing symbology of an entire layer (ArcGIS)](https://harvard.zoom.us/rec/play/aK8Cbw2IPOXzIZfhPOlrGOmQ_5BwvXvUErgbqJ2h5i41QBDA-ETqCyL8k4HHxEDcckH6g3RP0o8vfA.-BS6HcP9XkFX8zxb?autoplay=true&startTime=1660832780000){target="_blank"}
* [Changing symbology of an entire layer (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-layer-symbology.html){target="_blank"}
* [Choosing a basemap (ArcGIS)](https://harvard.zoom.us/rec/play/1XuBLT-4CY1-Lbgvrs-t6VXYJsciyfiMsH3XzPBaCi1v-3qGZnT9CpNbeBC8BRaBJ6LdvNgVwjj1E3Dw.H05muzT3eZDJloyL?autoplay=true&startTime=1660837402000){target="_blank"}
* [Choosing a basemap (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-basemap.html){target="_blank"}
* [Creating a legend for multiple layers in a ggplot map (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-multilayer-legend.html){target="_blank"}
* Creating an interactive map (R - leaflet)
* [Customizing a ggplot theme (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-custom-themes.html){target="_blank"}
* [Customizing a legend (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-custom-legend.html){target="_blank"}
* [Defining the bounding box for a ggplot map (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-bounding-box.html){target="_blank"}
* [Downloading ArcGIS extension for Illustrator](https://www.esri.com/en-us/arcgis/products/maps-for-adobecc/overview){target="_blank"}
* [Editing a ggplot map/chart further in Adobe Illustrator (Illustrator)](https://asngrads.com/2019/11/24/enhancing-figures-ggplot2-to-adobe-illustrator/){target="_blank"}
* [Editing a map file in Illustrator (ArcGIS)](https://harvard.zoom.us/rec/play/W7I2et1sT0FbmuastV4wlhsS8p-9A5tiD8yPlUkg9B2ZoJDPmzgyLG_4EIo0oQH-bh8Z8dCBs6silUZv.LKz43-ZilnmFeBzs?autoplay=true&startTime=1660840146000){target="_blank"}
* [Export a map layout (ArcGIS)](https://harvard.zoom.us/rec/play/Rqo-d93n-3c4Y69bjzfVccNEN0OsM91rk4yRrhlUMpI7qTi0Lm9ksX8xRSEPcWNVkbtT5oKdtgMgzLIU.eUdWhqVKWKeTfNMo?autoplay=true&startTime=1660839155000){target="_blank"}
* [Fun with color palettes (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-colors.html){target="_blank"}
* [Fun with fonts (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-fonts.html){target="_blank"}
* [Opening an AIX file in Illustrator (ArcGIS)](https://harvard.zoom.us/rec/play/6pVTkEkqeU9H60mD5zkCRLaxfcirCeBAAmh8DDUprxNVDR6zy4M18IgLkOrgsT6iEJRBBGSbOyJ94VQB._wEDohMqh1WEykl-?autoplay=true&startTime=1660839734000){target="_blank"}
* [Saving a ggplot map as a PDF or image (R)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-save-ggplot.html){target="_blank"}
* [Selecting a map projection (ArcGIS)](https://harvard.zoom.us/rec/play/evJJKuzp7TDTvzRyBdMfbYfcrJl4bk7zz77gGGEy7caGnVxUuRnSIgUjlubM18-XHfqNH_saovRKLJtn.aJkRu-YWDglMS_Xj?autoplay=true&startTime=1660837959000){target="_blank"}
* [Selecting a map projection (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-ggplot-proj.html){target="_blank"}

# Other resources

* [r-spatial tutorial on ggplot maps](https://r-spatial.org/r/2018/10/25/ggplot2-sf.html){target="_blank"}
* [Spatial Reference (for finding map projections)](https://spatialreference.org/){target="_blank"}
* [PROJ projections list](https://proj.org/operations/projections/index.html){target="_blank"}

# Examples

Each of the examples below fully meets the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/Examples/slo-leaflet.html){target="_blank"}
* [Example 2](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/Examples/Africa-a2.pdf){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Inclusion of multiple layers",
                          "The map is informative",
                          "The map is attractive",
                          "Source attribution",
                          "Description of audience",
                          "Description of argument/message",
                          "Total"),
                 `No credit` = c(rep("Missing", 8), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "",
                                   "Minimal apparent effort",
                                   "Minimal apparent effort",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described", 
                                   ""),
                 `Full credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "Included",
                                   "Evident effort",
                                   "Evident effort",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 4, 4, 4, 1, 4, 4, 25))

kable(rubric)
```
