---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Submit a map with showing the data you described in 
Assignment 1, with each dataset as a separate layer. 

Your map should be attractive and informative (inset maps, north arrows, and 
scale bars are very helpful in making a map informative, but I'm not 
dogmatic about including them if your audience doesn't need them). It should 
clearly indicate the source of each layer in your map. It should
also include a couple of sentences describing the audience for your map and
the argument you are making or the message you are conveying to them with
your map.

On your submission, list the name(s) of at least one student who was 
especially helpful you you in completing this assignment, and describe 
how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this assignment, but all 
of them are relevant and you will probably use several of them. They are
listed alphabetically.

* Adding a north arrow to a ggplot map (R)
* Adding a scale bar to a ggplot map (R)
* Annotating a ggplot map (R)
* Applying a ggplot theme (R)
* Changing symbology based on a variable in ggplot (R)
* Changing symbology of an entire layer in ggplot (R)
* Creating a legend for multiple layers in a ggplot map (R)
* Creating a legend for one layer in a ggplot map (R)
* Creating an interactive map with leaflet (R)
* Customizing a ggplot theme (R)
* Defining the bounding box for a ggplot map (R)
* Editing a ggplot map further in Adobe Illustrator (Illustrator)
* [Fun with color palettes (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-colors.html){target="_blank"}
* Fun with fonts (R)
* Saving a ggplot map as a PDF (R)
* [Selecting a projection for a ggplot map (R)](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-ggplot-proj.html){target="_blank"}

# Other resources

* [r-spatial tutorial on ggplot maps](https://r-spatial.org/r/2018/10/25/ggplot2-sf.html){target="_blank"}
* [Spatial Reference (for finding map projections)](https://spatialreference.org/){target="_blank"}
* [PROJ projections list](https://proj.org/operations/projections/index.html){target="_blank"}

# Examples

Each of the examples below fully meets the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/Examples/slo-leaflet.html){target="_blank"}
* [Example 2](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/Examples/Africa-a2.pdf){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Inclusion of multiple layers",
                          "The map is informative",
                          "The map is attractive",
                          "Source attribution",
                          "Description of audience",
                          "Description of argument/message",
                          "Total"),
                 `No credit` = c(rep("Missing", 8), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "",
                                   "Minimal apparent effort",
                                   "Minimal apparent effort",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described", 
                                   ""),
                 `Full credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "Included",
                                   "Evident effort",
                                   "Evident effort",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 4, 4, 4, 1, 4, 4, 25))

kable(rubric)
```
