---
title: "Saving a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

By default, your leaflet map initially be set to show all the data
in all the vector layers it includes. You may want to set the initial
view to something else.

# Required packages

This example uses the `leaflet` package and the `tidyverse` package. 

```{r, results='hide', message=FALSE}
library(leaflet)
library(tidyverse)
```

# Setting the initial view and zoom level

You can use the `setView()` function to center your map on a particular
point by specifying latitude/longitude coordinates. You can also set the initial 
zoom level.

Here's an example where I have no data in my map (just a basemap), so I need
to set the view manually. Here's a map that's centered on my old high school
(Go Bulldogs!).

```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = -111.659566, lat = 40.248393, zoom = 18)
```

Higher values for the zoom level are more "zoomed in". Lower values are more "zoomed out."

```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = -111.659566, lat = 40.248393, zoom = 10)
```


A zoom value of `zoom = 1` will show the whole world (still centered on my high school 
in this example).

```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = -111.659566, lat = 40.248393, zoom = 1)
```