---
title: "Saving images and PDFs of ggplots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Once you've put all that work into creating a beautiful map in R, you may 
want to save it for use outside of RStudio. 

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
```

# Loading some sample data

Let's start with this nice map I created that shows bus stop and trails in San 
Luis Opispo. It includes some text annotation as well.

```{r, include=FALSE}
city_trails <- here::here("a2-files",
                    "Examples",
               "Trail.geojson") %>%
  sf::st_read() %>%
  rename(TrailName = Trail_Name,
         Length = SLength) %>%
  mutate(LengthUnits = "MILES") %>%
  select(TrailName, Length, LengthUnits)

slo_stops <- here::here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  sf::st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  sf::st_set_crs("WGS84")

boundary <- tigris::places(state = "CA") %>%
  filter(NAME == "San Luis Obispo")

slo_map <- ggplot(city_trails) +
  ggspatial::annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trail")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stop")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary"),
          key_glyph = "path") +
  scale_color_manual(name = "Legend",
                     values = c("cornflowerblue", "black", "darkgreen"),
                     guide = guide_legend(override.aes = list(linetype = 
                                                         c("blank", "solid", "solid"),
                                                       shape = c(16, NA, NA)))) +
  theme(legend.background = element_rect(fill = NA),
        legend.key = element_rect(fill = NA),
        axis.title = element_blank()) +
  ggtitle(label = "Trails and Transit",
          subtitle = "Locations of city managed trails and bus stops\nin San Luis Obispo")
```

```{r, fig.height=7, message=FALSE}
slo_map
```

# Save as a pdf

The `ggsave()` function makes it really easy to save figures created with `ggplot()`.

This will save the most recent thing you plotted to your current working
directory, with the filename you specify (and it will guess the file format from
the file extension you give it)

```{r}
ggsave("slo_map.pdf")
```

This is a little risky. It is safer to specify in your code what exactly
it is you want to save and where you want to save it.

```{r}
ggsave(filename = here("a2-files",
                       "slo_map.pdf"),
       plot = slo_map)
```

You can also specify the size of the image you want to save.
If you want to save the file as a full page (8.5 inches by 11 inches), 
you would specify that like this ([here is the result](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/slo_map.pdf){target="_blank"}.

```{r}
ggsave(filename = here("a2-files",
                       "slo_map.pdf"),
       plot = slo_map,
       height = 11,
       width = 8.5,
       units = "in")
```

# Save as an image

If you want to save your file as an image, you can also specify the image 
resolution ([here is the result](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/slo_map.jpg){target="_blank"}.

```{r}
ggsave(filename = here("a2-files",
                       "slo_map.jpg"),
       plot = slo_map,
       height = 11,
       width = 8.5,
       units = "in",
       dpi = 300)
```