---
title: "Create a legend with multiple layers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating a single legend that includes the symbology for
multiple layers is trickier than in should be in ggplot,
but it's doable with a couple little hacks.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggspatial)
```

# Sample data

```{r, include=FALSE}
city_trails <- here::here("a2-files",
                    "Examples",
               "Trail.geojson") %>%
  sf::st_read() %>%
  rename(TrailName = Trail_Name,
         Length = SLength) %>%
  mutate(LengthUnits = "MILES") %>%
  select(TrailName, Length, LengthUnits)

slo_stops <- here::here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  sf::st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  sf::st_set_crs("WGS84")

boundary <- tigris::places(state = "CA") %>%
  filter(NAME == "San Luis Obispo")
```

Let's make a nice map that shows bus stop, trails, and the 
city boundary for San Luis Opispo, California.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          color = "darkgreen") +
  geom_sf(data = slo_stops,
          color = "cornflowerblue") +
  geom_sf(data = boundary,
          fill = NA,
          color = "black") +
  theme_void() 

slo_map
```

It would be helpful to have a legend showing what all those little 
dots and lines mean. `ggplot` will automatically generate a legend
when you set symbology within the `aes()` function (which you'd 
generally only do if you wanted to vary the symbology based on the
value of a variable), but not if you specify the color outside the
`aes()` function (which you'd generally do if you wanted to set the
symbology for the entire layer).

# Creating a legend

You can get a legend by setting the symbology by setting the name that 
should appear in the legend within the `aes()` function.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trails")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stops")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary")) +
  theme_void() 

slo_map
```

This will just assign default colors, but you can customize those
(and change the legend title) using `scale_color_manual()`.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trails")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stops")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary")) +
  scale_color_manual(name = "Legend",
                     values = c("cornflowerblue", "black", "darkgreen")) +
  theme_void() 

slo_map
```

Now the legend has all the right colors, but it's representing all three layers
as a combination of points, lines, and polygons. We can use the `override.aes`
parameter within the `guide_legend()` function to make it so the legend only
shows a point for bus stops. We give a list of shapes to display for points
in each legend item (in the same order they appear in the legend). Bus stops
are first, and we'll use shape 16 (see the [tutorial on symbology here](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-layer-symbology.html)
for alternative point shapes) for that first one, then we'll use NA values to indicate that 
the legend shouldn't display any point for the other two.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trails")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stops")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary")) +
  scale_color_manual(name = "Legend",
                     values = c("cornflowerblue", "black", "darkgreen"),
                     guide = guide_legend(override.aes = 
                                            list(shape = c(16, NA, NA)))) +
  theme_void() 

slo_map
```

We can also use override.aes to stop the box and lines from showing
for the bus stops by setting the linetype to "blank" for that first
item.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trails")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stops")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary")) +
  scale_color_manual(name = "Legend",
                     values = c("cornflowerblue", "black", "darkgreen"),
                     guide = guide_legend(override.aes = 
                                            list(shape = c(16, NA, NA),
                                                 linetype = c("blank", 
                                                              "solid", 
                                                              "solid")))) +
  theme_void() 

slo_map
```

The only remaining problem is that the legend wants to represent a polygon's
border color as a box, but it wants to represent the color of a line layer
as a single line. We can force the legend to represent them the same way
by setting the `key_glyph` value in the `geom_sf()` function that draws the
polygon.

```{r, message=FALSE}
slo_map <- ggplot(city_trails) +
  annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trails")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stops")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary"),
          key_glyph = draw_key_path) +
  scale_color_manual(name = "Legend",
                     values = c("cornflowerblue", "black", "darkgreen"),
                     guide = guide_legend(override.aes = 
                                            list(linetype = c("blank", 
                                                              "solid",
                                                              "solid"),
                                                 shape = c(16, NA, NA)))) +
  theme_void() 

slo_map
```

