---
title: "Saving a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You should save your leaflet map as an html file that's viewable in a web browser.

# Save your map as knitted markdown file

If you're working in RMarkdown, you can just knit the file to generate an html file that includes
your map. You can use markdown text to add explanatory text before or after your map. In the
heading to any code chunks, replace `{r}` with `{r, include=FALSE}` so that the code doesn't
show up in the knitted file.

You can also apply a theme to your knitted file by editing the YAML header (the first few lines
of your RMarkdown file between the `---` lines). 
[Here is a gallery of some available themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/){target="_blank"}. For example, if
I wanted to apply the `cosmo` theme, I would replace the default YAML header with

`---`

    title: "My cool map"
    output: 
      html_document:
            theme: cosmo
            
`---`

# Save your map as a full-page html file

You can also save your map separately from your Rmd file using the 
`saveWidget()` function in the `htmlwidgets` package.

```{r, results='hide', message=FALSE}
library(leaflet)
library(htmlwidgets)
library(here)

my_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = -120.6592, lat = 35.28190, zoom = 11)

saveWidget(my_map, file = here("a2-files", "leaflet-map.html"))
```

