---
title: "Customizing a ggplot theme"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The display of everything on your map that isn't data can 
is controlled by the theme. You have a lot of flexibility here.
I'm just going to demonstrate a few things to get you started.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Sample data

```{r, include=FALSE}
countries <- rnaturalearth::ne_countries(continent = "Africa",
                          returnclass = "sf")
```

I have a dataframe called `countries` with national boundaries in
Africa. I can use `ggplot()` and `geom_sf()` to make a quick 
chloropleth map with variations in color to represent the estimated
GDP. Here is a map using `theme_void()` as a starting point.

```{r}
gdp_map <- ggplot(countries) +
  geom_sf(aes(fill = gdp_md_est)) +
  ggtitle(label = "Sample map of Africa",
          subtitle = "Variation in GDP") +
  theme_void()

gdp_map
```

This doesn't look too bad, and all of the edits I make below will
make it much, much worse. But there may be a situation in which
knowing how to make one or more of these edits will allow you to improve
a map.

# Plot background

I can set the color and outline for the plot background.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = "gray",
                                       linetype = "dashed"))
```

# Panel background

The panel background is the rectangle around your data (not including 
titles and legends), and you can 
customize it the same way.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"))
```

# Plot margins

You can also set the margins around the outside of the plot with
`plot.margin`. `t` is for top, `r` is for right, `b` is for bottom,
and `l` is for left.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"))
```

# Legend

You can place the legend to the right, left, top, or bottom of the
map.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = "bottom")
```

You can leave the legend off by setting `legend.position = "none"`.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = "none")
```

Or you can place it over the map by specifying coordinates (relative to 
the map panel). Placing it at `c(0.5, 0.5)` will place it right in the 
middle of your map (which you probably don't want).

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.5, 0.5))
```

Here's a better placement.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.2, 0.25))
```

You can also specify the legend background:

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.2, 0.25),
        legend.background = element_rect(fill = "yellow",
                                         color = NA))
```

The `legend.margin` parameter works the same way as `plot.margin`.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.2, 0.25),
        legend.background = element_rect(fill = "yellow",
                                         color = NA),
        legend.margin = margin(t = 6, 
                             r = 6, 
                             b = 6, 
                             l = 6, unit = "pt"))
```

# Axes

The `theme_void()` theme we used as a starting point has no axis
text, but you can add it back in.

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.2, 0.25),
        legend.background = element_rect(fill = "yellow",
                                         color = NA),
        legend.margin = margin(t = 6, 
                             r = 6, 
                             b = 6, 
                             l = 6, unit = "pt"),
        axis.text = element_text())
```

# Gridlines

And you can also customize grid lines

```{r}
gdp_map +
  theme(plot.background = element_rect(fill = "cornsilk",
                                       color = NA),
        panel.background = element_rect(fill = "cadetblue",
                                        color = "gray",
                                        linetype = "dotted"),
        plot.margin = margin(t = 10, 
                             r = 10, 
                             b = 10, 
                             l = 10, unit = "pt"),
        legend.position = c(0.2, 0.25),
        legend.background = element_rect(fill = "yellow",
                                         color = NA),
        legend.margin = margin(t = 6, 
                             r = 6, 
                             b = 6, 
                             l = 6, unit = "pt"),
        axis.text = element_text(),
        panel.grid = element_line(color = "pink",
                                  linetype = "dotted"))
```