---
title: "Adding data to a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Leaflet maps are generally going to be most useful if you include 
a basemap to orient your audience.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
```

# Loading some sample data

I'll load some data with bus stop locations in San Luis Obispo, California:

```{r, message=FALSE, results='hide'}
slo_stops <- here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  st_set_crs("WGS84")
```

And display them on a Leaflet map.

```{r}
leaflet() %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   fillOpacity = 0.6,
                   stroke = FALSE)
```

Not very useful, is it? Normally, you'll want some kind of basemap
behind your data so your viewer knows what they're looking at.

# Adding the default OpenStreetMap basemap

The `addTiles()` function will put a default basemap behind your data.

```{r}
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   fillOpacity = 0.6,
                   stroke = FALSE)
```

# Adding other provider tiles

The `addProviderTiles()` function gives you a few more options. You can preview them
[here](https://leaflet-extras.github.io/leaflet-providers/preview/){target="_blank"}.

Here are few examples:

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   fillOpacity = 0.6,
                   stroke = FALSE)
```

```{r}
leaflet() %>%
  addProviderTiles("CartoDB.Voyager") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   fillOpacity = 0.6,
                   stroke = FALSE)
```

```{r}
leaflet() %>%
  addProviderTiles("Esri.NatGeoWorldMap") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   fillOpacity = 0.6,
                   stroke = FALSE)
```