---
title: "Setting highlight options"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The nice thing about Leaflet maps is their interactivity.
Highlight options will change how a line or polygon layer 
is displayed when you hover over it.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
library(tidycensus)
```

# Loading some sample data

I'll load some data with census tracts in San Luis Obispo County, California:

```{r, message=FALSE, results='hide'}
tracts <- get_acs(geography = "tract", 
                  variables = "B06011_001", 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) %>%
  st_transform("WGS84")
```

And display them on a Leaflet map.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1)
```

# Setting highlight options

The `highlightOptions` parameter with the `highlightOptions()` function let
you set a new line weight, line color, fill color, and fill opacity for when 
the user hovers over a feature.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"))
```

This will probably make the user expect something to happen when they click on the feature, so you 
might want to also set up a [label or pop-up](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-leaflet-labels.html){target="_blank"}
if you are using the highlight options.