---
title: "Applying symbology uniformly for a layer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can control the way layers are displayed in a ggplot map.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Sample data

```{r, include=FALSE}
library(sf)

countries <- rnaturalearth::ne_countries(continent = "Africa",
                          returnclass = "sf")

airport_flows <- here::here("a2-files",
                            "wb_airport_flows") %>%
  sf::st_read() 

airport_arrivals <- airport_flows %>%
  sf::st_drop_geometry() %>%
  rename(iata_code = Dest) %>%
  group_by(iata_code) %>%
  summarise(arr_seats = sum(TotalSeats))

airports <- here::here("a1-files",
                 "sample-files",
                 "airports") %>%
  sf::st_read() %>%
  sf::st_filter(countries) %>%
  left_join(airport_arrivals) %>%
  filter(!is.na(arr_seats))

within_africa_flows <- airport_flows %>%
  filter(Orig %in% airports$iata_code &
           Dest %in% airports$iata_code) %>%
  mutate(flow_category = case_when(TotalSeats > 100000 ~ "> 100,000",
                                    TotalSeats < 25000 ~ "< 25000",
                                    TRUE ~ "25,000 to 100,000"))
```

Here is a map of African countries and airports, with lines representing the
flights between pairs of airports.

```{r}
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = within_africa_flows) +
  geom_sf(data = airports) +
  theme_bw()
```

# Line symbology

I can set line weight (`size = `), color (`color = ` or `colour =`), 
type (`lty =`), and transparency (`alpha =`).

Options for line types are:

* `blank`
* `solid`
* `dashed`
* `dotted`
* `dotdash`
* `longdash`
* `twodash`

```{r}
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = within_africa_flows,
          color = "cornflowerblue",
          size = 0.5,
          lty = "dotted",
          alpha = 0.3) +
  geom_sf(data = airports) +
  theme_bw()
```

# Point symbology

You can specify point shapes with any of the following numbers:

```{r, echo=FALSE}
shapes <- tibble(key = c(paste0("0", seq(0,9,by=1)), seq(10,25, by=1)),
                 label = as.character(seq(0, 25, by=1)),
                 x =c(rep(seq(1,6, by=1),4),1, 2),
                 y = c(rep(6, 6), rep(5, 6),
                       rep(4, 6), rep(3, 6),
                       rep(2,2)))

ggplot(shapes) +
  geom_point(aes(x = x, 
                 y = y,
                 shape = key),
             size = 5,
             fill = "pink") +
  geom_text(aes(x = x, 
                 y = y,
                 label = label),
             nudge_y = -0.25) +
  scale_shape_manual(values = seq(0, 25, by=1)) +
  theme_void() +
  theme(legend.position = "none")
```

```{r}
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = within_africa_flows,
          color = "cornflowerblue",
          size = 0.5,
          lty = "dotted",
          alpha = 0.3) +
  geom_sf(data = airports,
          shape = 3) +
  theme_bw()
```

You can also specify color, size, and transparency.

```{r}
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = within_africa_flows,
          color = "cornflowerblue",
          size = 0.5,
          lty = "dotted",
          alpha = 0.3) +
  geom_sf(data = airports,
          shape = 16,
          size = 4,
          alpha = 0.3,
          color = "red") +
  theme_bw()
```

For shapes 21 through 25, you can specify a color and a fill color:

```{r}
ggplot() +
  geom_sf(data = countries) +
  geom_sf(data = within_africa_flows,
          color = "cornflowerblue",
          size = 0.5,
          lty = "dotted",
          alpha = 0.3) +
  geom_sf(data = airports,
          shape = 21,
          size = 4,
          alpha = 0.3,
          color = "red",
          fill = "blue") +
  theme_bw()
```

# Polygon symbology

For polygons, you can specify the fill color, the fill transparency,
the outline color, the outline weight, and the outline line type.

```{r}
ggplot(countries) +
  geom_sf(size = 1.5,
          lty = "dotted",
          color = "pink",
          fill = "yellow",
          alpha = 0.2) +
  theme_bw()
```

If you set `color = NA` you can leave off the outlines.

```{r}
ggplot(countries) +
  geom_sf(color = NA,
          fill = "yellow",
          alpha = 0.2) +
  theme_bw()
```
