---
title: "Basemaps in ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `ggspatial` package has a few different basemaps you can choose from.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggspatial)
```

# Sample data

Here's a map of bus stop in San Luis Obispo, California.

```{r, include=FALSE}
slo_stops <- here::here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  sf::st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  sf::st_set_crs("WGS84")

slo_streets <- here::here("a2-files",
                  "Examples",
                  "Streets.geojson") %>%
  sf::st_read()
```

```{r}
ggplot(slo_stops) +
  geom_sf() +
  theme_void() 
```

This is totally meaningless. A basemap would help my audience orient themselves.

# A vector layer as a basemap

I want to note that you can often get away without a basemap. For example, I could add a 
layer of roads, and that might be enough to orient my reader, especially if they are 
familiar with the area. Less is more, and all the detail that comes with a basemap
image might only serve to clutter up your map.

```{r}
ggplot(slo_stops) +
  geom_sf(data = slo_streets,
          color = "gray") +
  geom_sf() +
  theme_void() 
```

# An image basemap

The `ggspatial` package can also add basemaps to your map using the
`annotation_map_tile()` function. Here is is with the default basemap
from OpenStreetMaps.

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile() +
  geom_sf() +
  theme_void() 
```

The image looks a little blurry because the default is to give you a lower-resolution
image to that the map will load faster. You can fix this by setting `zoomin = 0`.

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(zoomin = 0) +
  geom_sf() +
  theme_void() 
```

# Alternative basemaps

You can select a different basemap using `type = `

Here are some options.

`hotstyle`

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(type = "hotstyle",
                      zoomin = 0) +
  geom_sf() +
  theme_void() 
```

`stamenbw`

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(type = "stamenbw",
                      zoomin = 0) +
  geom_sf() +
  theme_void() 
```

`stamenwatercolor`

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(type = "stamenwatercolor",
                      zoomin = 0) +
  geom_sf() +
  theme_void() 
```


`cartodark`

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(type = "cartodark",
                      zoomin = 0) +
  geom_sf(color = "yellow") +
  theme_void() 
```

`cartolight`

```{r, message=FALSE}
ggplot(slo_stops) +
  annotation_map_tile(type = "cartolight",
                      zoomin = 0) +
  geom_sf() +
  theme_void() 
```

There are some other cool ones that requre a (free!) API key, but I have not
figured out how to pass an API key to the `annotation_map_tile()` function.
Let me know if you figure out an easy approach!