---
title: "Adding a north arrow to a map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are those who might insist that every map needs a north arrow and 
a scale bar. I would argue that a north arrow is necessary if it's 
useful to the audience. If you are creating a map of South America, and
your audience is familiar enough with South American to recognize its shape,
you may not need a north arrow. 

For other maps, a north arrow is helpful.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(ggspatial)
```

# Loading some sample data

```{r, include=FALSE}
edmonton <- here::here("a2-files",
                 "Examples",
                 "edmonton.geojson") %>%
  st_read() %>%
  filter(grepl("Arterial", class_de))
```

I'll demonstrate with a map of arterial streets in Edmonton, Alberta 
(data are from Open Street Map via the `osmextract` package).

```{r}
edmonton_map <- ggplot(edmonton) +
  geom_sf() +
  theme_void()

edmonton_map
```

# Add a north arrow

The `ggspatial` package has a useful function called `annotation_north_arrow()`
that will add a north arrow to your map.

```{r}
edmonton_map +
  annotation_north_arrow() 
```

# Place the north arrow

By default, the north arrow is in the bottom left corner, but you can
specify which corner of the map to place it using the `location` argument.

* `tr` means "top right".
* `tl` means "top left".
* `br` means "bottom right".
* `bl` means "bottom left"

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl")
```

# Select a north arrow style

You have four different option for north arrow styles:

`north_arrow_orienteering` (this is the default)

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_orienteering())
```

`north_arrow_fancy_orienteering`

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_fancy_orienteering())
```

`north_arrow_minimal`

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal())
```

`north_arrow_nautical`

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_nautical())
```

# Customizing a north arrow

You can customize each of the above styles a little bit using the 
following parameters:

* line_width
* line_col
* fill (must be two colors, or one color for north_arrow_minimal)
* text_size
* text_face (e.g. `"bold"`, or `italic"`)
* text_family (see [this page for more on fonts](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-fonts.html){target="_blank"})
* text_col (must be two colors, except for north_arrow_minimal)
* text_angle (not available for north_arrow_minimal)

`north_arrow_minimal`

```{r}
edmonton_map +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_nautical(line_width = 2,
                                                     line_col = "gray",
                                                     fill = c("lightblue", "pink"),
                                                     text_size = 18,
                                                     text_face = "italic",
                                                     text_family = "serif",
                                                     text_col = "pink",
                                                     text_angle = 90))
```