---
title: "Adding labels or popups to a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Labels and popups can make a leaflet map more informative.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
library(tidycensus)
library(htmltools)
```

# Loading some sample data

I'll load some data with census tracts in San Luis Obispo County, California:

```{r, message=FALSE, results='hide'}
tracts <- get_acs(geography = "tract", 
                  variables = "B06011_001", 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) %>%
  st_transform("WGS84")
```

And display them on a Leaflet map. I've set highlight options so that a tract turns
yellow when you hover over it, but nothing else happens.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"))
```

# Add labels

I can add a label so some text will appear when I highlight a census tract. Here is how
I would display the value of the "NAME" variable.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"),
              label = ~NAME)
```

# Add popups

In Leaflet, labels appear when you hover over a feature, and popups only appear
when you click on a feature. 

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"),
              popup = ~NAME)
```

I can also customize and format my popup text by creating a new variable in the data.

```{r}
tracts <- tracts %>%
  mutate(popup_text = paste0("Median income: $", 
                             prettyNum(B06011_001E, big.mark = ",")))

leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"),
              popup = ~popup_text)
```

You can add some html tags to the popup text as well.

* `<br/>` will insert a line break.
* Text between `<b>` and `</b>` will be in bold.

I don't know why, but html tags don't seem to work for labels, only
for popups.

```{r}
tracts <- tracts %>%
  mutate(popup_text = paste0(NAME, 
                             "<br/><b>Estimated median income:</b> $", 
                             prettyNum(B06011_001E, big.mark = ",")))

leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1,
              highlightOptions = highlightOptions(weight = 2,
                                                  fillOpacity = 0.6,
                                                  color = "yellow",
                                                  fillColor = "cornsilk"),
              popup = ~popup_text)
```
