---
title: "Adding data to a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `leaflet` package in R helps you to create interactive, web-based maps.
This page (https://rstudio.github.io/leaflet/) gives a nice overview. I'll give
some brief examples here.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
library(tigris)
```

# Loading some sample data

Let's start by loading some data with the trails and bus stops, and  in San Luis Obispo, 
California

```{r, message=FALSE, results='hide'}
city_trails <- here("a2-files",
                    "Examples",
                    "Trail.geojson") %>%
  st_read()

slo_stops <- here::here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  st_set_crs("WGS84")

boundary <- places(state = "CA") %>%
  filter(NAME == "San Luis Obispo") %>%
  st_transform("WGS84")
```

# Point data

I set up a leaflet map using the `leaflet()` function. Then, after a pipe (`%>%`), 
I can add point data using the `addMarkers()`. For 

```{r}
leaflet() %>%
  addMarkers(data = slo_stops)
```

As an alternative, I can also display point data as little dots by using `addCircleMarkers()` instead
of `addMarkers()`.

```{r}
leaflet() %>%
  addCircleMarkers(data = slo_stops)
```

With circle markers, you can specify the radius, outline color/opacity/weight, and fill color/opacity. 

```{r}
leaflet() %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "lightgreen",
                   fillOpacity = 0.6,
                   color = "darkgreen",
                   weight = 1,
                   opacity = 1)
```

And you can also leave off the outline by setting `stroke = FALSE`,

```{r}
leaflet() %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   stroke = FALSE)
```

# Polygon data

You can add polygon data to your leaflet map using `addPolygons()`.

```{r}
leaflet() %>%
  addPolygons(data = boundary) %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   stroke = FALSE) 
```

And I can customize the appearance of the polygon in the same way I would customize the circle
markers.

```{r}
leaflet() %>%
  addPolygons(data = boundary,
              fillOpacity = 0,
              weight = 1.5,
              color = "black") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   stroke = FALSE) 
```

# Line data

You can also add a line layer.

```{r}
leaflet() %>%
  addPolygons(data = boundary,
              fillOpacity = 0,
              weight = 1.5,
              color = "black") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   stroke = FALSE) %>%
  addPolylines(data = city_trails)
```

Again, you can mess around with line weight, opacity, and color.

```{r}
leaflet() %>%
  addPolygons(data = boundary,
              fillOpacity = 0,
              weight = 1.5,
              color = "black") %>%
  addCircleMarkers(data = slo_stops,
                   radius = 4,
                   fillColor = "darkgreen",
                   stroke = FALSE) %>%
  addPolylines(data = city_trails,
               weight = 2,
               color = "brown",
               opacity = 1)
```
