---
title: "Themes in ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The package `ggthemes` has several predefined themes you can
customize a ggplot graph, chart, or map. 

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
```

# Sample data

```{r, include=FALSE}
countries <- rnaturalearth::ne_countries(continent = "Africa",
                          returnclass = "sf")
```

I have a dataframe called `countries` with national boundaries in
Africa. I can use `ggplot()` and `geom_sf()` to make a quick 
chloropleth map with variations in color to represent the estimated
GDP.

```{r}
gdp_map <- ggplot(countries) +
  geom_sf(aes(fill = gdp_md_est))

gdp_map
```

# Apply a theme

The map above uses the default theme for ggplot maps. It has a gray
background with white gridlines, and a legend on the righthand side 
of the plot. The `ggplot2` package comes with other options for themes,
such as `theme_bw()`:

```{r}
gdp_map +
  theme_bw()
```

`theme_dark()`:

```{r}
gdp_map +
  theme_dark()
```

Or `theme_minimal()`:

```{r}
gdp_map +
  theme_minimal()
```

For the most part, these themes were designed with scatterplots
in mind. More often than not, for a map, you just want a blank
background with no gridlines or axis labels. You can get that
with `theme_void()`:

```{r}
gdp_map +
  theme_void()
```

The `ggthemes` package has even more themes to choose from, like 
`theme_economist()`, which is meant to make a graph in the style of
*The Economist* magazine.

```{r}
gdp_map +
  theme_economist()
```

Or `theme_fivethirtyeight`, which is meant to make a graph look like the
ones published by Nate Silver's FiveThirtyEight website.

```{r}
gdp_map +
  theme_fivethirtyeight()
```

Again, most of these were designed for scatter plots and similar graphs. They 
might be less useful for maps. But one `ggthemes` theme that is useful for 
maps is `theme_map`. It looks a lot like `theme_void()`, but it places the
legend on top of the map, in the bottom left corner, and makes the legend text a 
little smaller.

```{r}
gdp_map +
  theme_map()
```
