---
title: "Annotate a map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's say I've created a nice map in ggplot, and now I want to add some 
text to it for context. I could easily just save my map as an image 
and add it to a layout in InDesign (or PowerPoint, or Word, for that 
matter), but you can add annotation within your code as well.

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
```

# Sample data

```{r, include=FALSE}
city_trails <- here::here("a2-files",
                    "Examples",
               "Trail.geojson") %>%
  st_read() %>%
  rename(TrailName = Trail_Name,
         Length = SLength) %>%
  mutate(LengthUnits = "MILES") %>%
  select(TrailName, Length, LengthUnits)

slo_stops <- here::here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
  st_set_crs("WGS84")

boundary <- tigris::places(state = "CA") %>%
  filter(NAME == "San Luis Obispo")
```

Let's start with this nice map I created that shows bus stop and trails in San 
Luis Opispo.

```{r, include=FALSE}
slo_map <- ggplot(city_trails) +
  ggspatial::annotation_map_tile(type = "hotstyle",
                                 zoomin = 0) +
  geom_sf(data = city_trails,
          aes(color = "Trail")) +
  geom_sf(data = slo_stops,
          aes(color = "Bus stop")) +
  geom_sf(data = boundary,
          fill = NA,
          aes(color = "City boundary"),
          key_glyph = "path") +
  scale_color_manual(name = "",
                     values = c("cornflowerblue", "black", "darkgreen"),
                     guide = guide_legend(override.aes = list(linetype = 
                                                         c("blank", "solid", "solid"),
                                                       shape = c(16, NA, NA)))) +
  ggthemes::theme_map() +
  ggspatial::annotation_north_arrow(location = "tr",
                         style = ggspatial::north_arrow_minimal()) +
  ggspatial::annotation_scale(location = "br") +
  theme(legend.background = element_rect(fill = NA),
        legend.key = element_rect(color = "white"))
```

```{r, fig.height=7, message=FALSE}
slo_map
```

# Adding a title

I can add a title and subtitle to the top of my map using the `ggtitle()`
function

```{r, fig.height=7, message=FALSE}
slo_map +
  ggtitle(label = "Trails and Transit",
          subtitle = "Locations of city managed trails and bus stops in San Luis Obispo")
```

# Placing annotation on the map

To place annotation text on the map itself, I'll need to specify coordinates
for where to place it.

## Adding x/y scales.

My map looks nice without any scales on the x and y axes. I'm using `theme_map()`, 
which leaves them off, but I'm going to temporarily put some back on just to have
a sense of the coordinate system so I can see where to place items on the map.

Looking at the map below, if I want to add some text to the top left corner 
of the map, I'd want the top left corner of the text block to be at 
about 35.36 degrees north and 120.73 degrees west.

```{r, fig.height=7, message=FALSE}
slo_map +
  ggtitle(label = "Trails and Transit",
          subtitle = "Locations of city managed trails and bus stops in San Luis Obispo") +
  theme(axis.text = element_text())
```

## Adding text

The `annotate()` function will add text to the map. I need to specify x and y 
coordinates to place the text. The `hjust`(horizonal justification)  and `vjust`
(vertical justification) arguments indicate what part of the block of text these
coordinate refer to:

* `hjust = 0` means the point you've indicated is the left side of the text
block (and the text will be left-justified).
* `hjust = 0.5` means the point you've indicated is the middle (horizontally) 
of the text block (and the text will be centered).
* `hjust = 1` means the point you've indicated is the the right side of the 
text block (and the text will be right-justified).
* `vjust = 0` means the point you've indicated is the bottom of the text block.
* `vjust = 0.5` means the point you've indicated is the midde (vertically) of 
the text block.
* `vjust = 1` means the point you've indicated is the top of the text block.

If you want your text to wrap across multiple lines, use `\n` within your
text string. This will insert a line break.

```{r, fig.height=7, message=FALSE}
slo_map +
  ggtitle(label = "Trails and Transit",
          subtitle = "Locations of city managed trails and bus stops in San Luis Obispo") +
  annotate(geom = "text",
           x = -120.73,
           y = 35.36,
           hjust = 0,
           vjust = 1,
           label = "This map shows that the areas\nof San Luis Obispo served by\ntrails are distinct from (and \nminimally overlapping with) the\nareas served by transit. Few trails\nare easily accessible by transit.")
```
