---
title: "Adding a legend to a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A legend or other annotation can make your leaflet map
more informative.

# Required packages

These examples uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
library(tidycensus)
library(htmltools)
```

# Loading some sample data

I'll load some data with census tracts in San Luis Obispo County, California:

```{r, message=FALSE, results='hide'}
tracts <- get_acs(geography = "tract", 
                  variables = c(median_income = "B06011_001"), 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) %>%
  st_transform("WGS84") %>%
  filter(!is.na(median_incomeE))
```

And display them on a Leaflet map, where the fill color of the map varies by median income. 

```{r}
income_palette = colorNumeric(palette = "RdYlBu", 
                              domain = tracts$median_incomeE)

leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              color = "black",
              fillOpacity = 0.7,
              weight = 0.3,
              fillColor = ~income_palette(median_incomeE))
```

Unlike ggplot, leaflet won't automatically add a legend to interpret the colors.
You'll need to [add that in yourself](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-leaflet-legend.html){target="_blank"}.

# Add a legend

I can add a legend to my map using the `addLegend()` function. I need to specify
which dataset the legend refers to (`data = tracts`), which color palette to use
(`pal = income_palette` - the one you defined previoulsy - 
[see how to define a color ramp here](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-leaflet-colors.html){target="_blank"}),
and which variable the color ramp is mapped to (`values = ~median_incomeE`).

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              color = "black",
              fillOpacity = 0.7,
              weight = 0.3,
              fillColor = ~income_palette(median_incomeE)) %>%
    addLegend(data = tracts,
              pal = income_palette, 
              values = ~median_incomeE)
```

# Format the legend

I can also specify the legend title and format the value labels a little by adding a dollar sign as a 
prefix.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              color = "black",
              fillOpacity = 0.7,
              weight = 0.3,
              fillColor = ~income_palette(median_incomeE)) %>%
    addLegend(data = tracts,
              pal = income_palette, 
              values = ~median_incomeE,
              title = "Estimated Median Income",
              labFormat = labelFormat(prefix = "$"))
```

# Position the legend

I can specify the position of the legend as one of the following:

* `"topright"`
* `"bottomright"`
* `"bottomleft"`
* `"topleft"`

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              color = "black",
              fillOpacity = 0.7,
              weight = 0.3,
              fillColor = ~income_palette(median_incomeE)) %>%
    addLegend(data = tracts,
              pal = income_palette, 
              values = ~median_incomeE,
              title = "Estimated Median Income",
              labFormat = labelFormat(prefix = "$"),
              position = "bottomleft")
```

