---
title: "Setting the map projection in ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When you're mapping a large area (like a continent or a large country) where the
geometry is stored as geographic coordinates, you'll essentially end up with a 
Mercator projection that distorts areas on the map.

This is less of a big deal for smaller areas like cities, and I guess it's not a 
big deal if you and your audience doesn't hate the Mercator projection (and I do hate 
it).

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
```

# Loading some sample data

```{r, message=FALSE, echo=FALSE, results='hide'}
provinces <- rnaturalearth::ne_states(country = "Canada", 
                                      returnclass = "sf")

world <- rnaturalearth::ne_countries(returnclass = "sf")
```

I'll demonstrate with map of the world and provincial boundaries in Canada.
Both have geometry stored in WGS 84 coordinates. 

Here is the world:

```{r}
wgs84_world <- ggplot(world) +
  geom_sf() +
  theme_minimal()

wgs84_world
```

And here are the Canadian provinces:

```{r}
wgs84_provinces <- ggplot(provinces) +
  geom_sf() +
  theme_minimal()

wgs84_provinces
```

The fidelity of these maps to actual areas and distances gets worse the further 
from the equator (north or south) you go. This may or may not be a problem
for you, but if you're using a Mercator projection, it should be because
you've decided that it's the best choice, not because you don't know how
to change it.

# Projections for world maps

The PROJ website has [a list of 
projections](https://proj.org/operations/projections/index.html){target="_blank"} 
you can browse through, many of which could be suitable for a map of the world.

On the page for a projection, find the proj-string and use it to set the CRS
for your map with the `coord_sf` function. 

Examples include [azimuthal equidistant](https://proj.org/operations/projections/aeqd.html){target="_blank"}:

```{r}
wgs84_world +
  coord_sf(crs = "+proj=aeqd") 
```

Or [equal earth](https://proj.org/operations/projections/eqearth.html){target="_blank"}:

```{r}
wgs84_world +
  coord_sf(crs = "+proj=eqearth") 
```

# PROJ strings

The examples above set the CRS with string starting with `+proj=`. This is
proj4 notation, and it's a common way to refer to a projection. 

proj4 notation lets you set several additional parameters. The most important
ones will often be to center your projection on a point near the middle of 
your map. If you don't set parameters to center your projection, your projection
will be centered on a point near the western part of Africa (zero degrees latitude
and zero degrees longitude). This will look wacky if you're mapping an area that's
far away from that point. For example, here is the azimuthal equidistant
projection applied to Canada.

```{r}
wgs84_provinces +
  coord_sf(crs = "+proj=aeqd") 
```

If you go 
back to the page for that projection (https://proj.org/operations/projections/aeqd.html),
scroll down to "Options" and you'll see that you can set parameters for `lat_0` and 
`lon_0`. You can use these to recenter the projection at 60 degrees north,
106 degrees west (about the middle of the northern border of Saskatchewan).

```{r}
wgs84_provinces +
  coord_sf(crs = "+proj=aeqd +lat_0=60 +lon_0=-106") 
```

Some projections that were created specifically for world maps have parameters
to center the map by longitude, but not by latitude. The Equal Earth projection
is one of these. 

Here is Canada in the Equal Earth projection centered on the prime meridian:

```{r}
wgs84_provinces +
  coord_sf(crs = "+proj=eqearth") 
```

And here it is centered at 106 degrees west:

```{r}
wgs84_provinces +
  coord_sf(crs = "+proj=eqearth +lon_0=-106") 
```

# Other proj4 examples

Feel free to experiment with different map projections for the area you are
mapping! 

Here are some reasonable ones to consider:

* **United States:** USA Contiguous Albers Equal Area Conic
`+proj=aea +lat_1=29.5 +lat_2=45.5 +lon_0=-96 +datum=NAD83 +units=m +no_defs`
    * `+proj=aea` indicates an Albers Equal Area projection (prioritizing preserving relative areas).
    * `+lat_1=29.5` indicates the projection plane intersects the ellipsoid at 29.5 degrees north.
    * `+lat_2=45.5` indicates the projection plane intersects the ellipsoid at 45.5 degrees north.
    * `+lon_0=-96` indicates the projection is centered at 96 degrees west.
    * `+datum=NAD83` indicates the projection is based on the NAD83 datum.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **Canada:** Statistics Canada Lambert: `+proj=lcc +lat_1=49 +lat_2=77 +lat_0=63.390675 +lon_0=-91.866667 +x_0=6200000 +y_0=3000000 +ellps=GRS80 +units=m +no_defs`
    * `+proj=lcc` indicates an Lambert Conformal Conic projection (prioritizing preserving angles and shapes).
    * `+lat_1=49` indicates the projection plane intersects the ellipsoid at 49 degrees north.
    * `+lat_2=77` indicates the projection plane intersects the ellipsoid at 77 degrees north.
    * `+lat_0=63.390675` indicates the projection is centered at 63.390675 degrees north.
    * `+lon_0=-91` indicates the projection is centered at 91.866667 degrees west.
    * `+x_0=6200000` means a value of 6,200,000 meters will be added to the map's x-coordinates 
    (this is to avoid negative values to to the west of the projection's center).
    * `+y_0=3000000` means a value of 3,000,000 meters will be added to the map's y-coordinates 
    (this is to avoid negative values to to the south of the projection's center).
    * `+ellps=GRS80` indicates the projection is based on the GRS 80 ellipsoid.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **Africa:** Africa Lambert Conformal Conic: `+proj=lcc +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +ellps=WGS84 +datum=WGS84 +units=m +no_defs`
    * `+proj=lcc` indicates an Lambert Conformal Conic projection (prioritizing preserving angles and shapes).
    * `+lat_1=20` indicates the projection plane intersects the ellipsoid at 20 degrees north.
    * `+lat_2=-23` indicates the projection plane intersects the ellipsoid at 23 degrees south.
    * `+lat_0=0` indicates the projection is centered at the equator.
    * `+lon_0=25` indicates the projection is centered at 25 degrees east.
    * `+ellps=WGS84` indicates the projection is based on the WGS 84 ellipsoid.
    * `+datum=WGS84` indicates the projection is based on the WGS 84 datum.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **South America:** South America Equidistant Conic: `+proj=eqdc +lon_0=-60 +lat_1=-5 +lat_2=-42 +ellps=aust_SA +units=m +no_defs`
    * `+proj=eqdc` indicates an Equidistant Conic projection (prioritizing preserving relative distances).
    * `+lat_1=-5` indicates the projection plane intersects the ellipsoid at 5 degrees south.
    * `+lat_2=-42` indicates the projection plane intersects the ellipsoid at 42 degrees south.
    * `+lon_0=-60` indicates the projection is centered at 60 degrees west.
    * `+ellps=aust_SA` indicates the projection is based on the Australian National and South American ellipsoid.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **Russia:** Albers Equal Area Russia: `+proj=aea +lat_1=50 +lat_2=70 +lon_0=100 +ellps=WGS84 +datum=WGS84 +units=m +no_defs`
    * `+proj=aea` indicates an Albers Equal Area projection (prioritizing preserving relative areas).
    * `+lat_1=50` indicates the projection plane intersects the ellipsoid at 50 degrees north.
    * `+lat_2=70` indicates the projection plane intersects the ellipsoid at 70 degrees north.
    * `+lon_0=100` indicates the projection is centered at 100 degrees east.
    * `+ellps=WGS84` indicates the projection is based on the WGS 84 ellipsoid.
    * `+datum=WGS84` indicates the projection is based on the WGS 84 datum.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **China:*** China Albers Equal Area Conic: `+proj=aea +lat_1=27 +lat_2=45 +lon_0=105 +ellps=WGS84 +datum=WGS84 +units=m +no_defs`
    * `+proj=aea` indicates an Albers Equal Area projection (prioritizing preserving relative areas).
    * `+lat_1=27` indicates the projection plane intersects the ellipsoid at 50 degrees north.
    * `+lat_2=45` indicates the projection plane intersects the ellipsoid at 70 degrees north.
    * `+lon_0=105` indicates the projection is centered at 105 degrees east.
    * `+ellps=WGS84` indicates the projection is based on the WGS 84 ellipsoid.
    * `+datum=WGS84` indicates the projection is based on the WGS 84 datum.
    * `units=m` indicates any distance measurements will be in meters.
    * `+no_defs` is just something to add at the end to keep defaults from doing something surprising.
* **Western Europe:** Lambert Azimuthal Equal Area: `+proj=laea +lat_0=52 +lon_0=20`
    * `+proj=aea` indicates an Lambert Azimuthal Equal Area projection (prioritizing preserving relative areas).
    * `+lat_0=52` indicates the projection is centered at 52 degrees north.
    * `+lon_0=20` indicates the projection is centered at 20 degrees east.
    
# Other ways to indicate map projections

The nice thing about proj4 notation is that a human can read and understand
the parameters it's applying. `sf` functions like `coord_sf` and `st_transform` 
will also understand other names for projections.

If you want to go with a straight Mercator projection, you can set the CRS 
for your map to `WGS84` or `NAD83` without dealing with a proj4 string. Just
set it as `crs = "WGS84`.

You can also use a 4-digit EPSG code for pre-defined map projections. Enter these
as numbers rather than strings (without quotations). You can find EPSG codes in 
a reference like [spatialrefernce.org](https://spatialreference.org/){target="_blank"}.