---
title: "Adding a scale bar to a map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are those who might insist that every map needs a north arrow and 
a scale bar. I would argue that a scale bar is necessary if it's 
useful to the audience. 

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(ggspatial)
```

# Loading some sample data

```{r, include=FALSE}
edmonton <- here::here("a2-files",
                 "Examples",
                 "edmonton.geojson") %>%
  st_read() %>%
  filter(grepl("Arterial", class_de))
```

I'll demonstrate with a map of arterial streets in Edmonton, Alberta 
(data are from Open Street Map via the `osmextract` package).

```{r}
edmonton_map <- ggplot(edmonton) +
  geom_sf() +
  theme_void()

edmonton_map
```

# Add a scale bar

The `ggspatial` package has a useful function called `annotation_scale()`
that will add a scale bar to your map.

```{r}
edmonton_map +
  annotation_scale() 
```

# Place the scale bar

By default, the scale bar is in the bottom left corner, but you can
specify which corner of the map to place it using the `location` argument.

* `tr` means "top right".
* `tl` means "top left".
* `br` means "bottom right".
* `bl` means "bottom left"

```{r}
edmonton_map +
  annotation_scale(location = "br")
```

# Customize a scale bar

You have several parameters you can specify:

* `bar_cols`: Two colors for the alternating bands on the bar (for `style = "bar"`)
* `line_width`: The line weight
* `height`: The width of the scale bar
* `pad_x`: The distance between the scale bar and the side of the map
* `pad_y`: The distance between the scale bar and the bottom (or top) of 
the map
* `text_pad`: The margin around the scale bar text
* `text_cex`: The text size
* `text_face` (e.g. `"bold"`, or `italic"`)
* `text_family` (see [this page for more on fonts](https://c-voulgaris.github.io/VIS-2128-F22/a1-files/skill-fonts.html){target="_blank"})
* `tick_height`: Height of tick marks, as a proportion of the scale bare
height (for `style = "ticks"`)
* `line_col`: The line color
* `text_col`: The text color
* `width_hint`: The (suggested) proportion of the map's width the scale
bar should take up
* `unit_category`: Either "metric" or "imperial"
* `style`: Either "ticks" or "bar"

```{r}
edmonton_map +
  annotation_scale(location = "br",
                   bar_cols = c("darkolivegreen", "cornsilk"),
                   line_width = 2,
                   height = unit(12, "points"),
                   pad_x = unit(3, "points"),
                   pad_y = unit(3, "points"),
                   text_pad = unit(12, "points"),
                   text_cex = 1,
                   text_face = "bold",
                   text_family = "serif",
                   line_col = "burlywood",
                   text_col = "burlywood4",
                   width_hint = 0.5,
                   unit_category = "imperial",
                   style = "bar")
```

```{r}
edmonton_map +
  annotation_scale(location = "br",
                   line_width = 0.5,
                   height = unit(6, "points"),
                   pad_x = unit(12, "points"),
                   pad_y = unit(0, "points"),
                   text_pad = unit(6, "points"),
                   text_cex = 0.7,
                   text_face = "italic",
                   text_family = "mono",
                   tick_height = 0.5,
                   line_col = "blue",
                   text_col = "red",
                   width_hint = 0.33,
                   unit_category = "metric",
                   style = "ticks")
```