---
title: "Applying symbology based on a variable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If you've [varied the symbology of a layer based on the value
of a variable](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-variable-symbology.html){target="_blank"}, 
a legend will be created automatically. You can customize this legend. 

# Required packages

This example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Sample data

```{r, include=FALSE}
library(sf)

countries <- rnaturalearth::ne_countries(continent = "Africa",
                          returnclass = "sf")

airport_flows <- here::here("a2-files",
                            "wb_airport_flows") %>%
  sf::st_read() 

airport_arrivals <- airport_flows %>%
  sf::st_drop_geometry() %>%
  rename(iata_code = Dest) %>%
  group_by(iata_code) %>%
  summarise(arr_seats = sum(TotalSeats))

airports <- here::here("a1-files",
                 "sample-files",
                 "airports") %>%
  sf::st_read() %>%
  sf::st_filter(countries) %>%
  left_join(airport_arrivals) %>%
  filter(!is.na(arr_seats))

```

Here is a map of African countries and airports. The point shape indicates the airport
type and the countries' fill colors indicate the estimated GDP.

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  theme_void()
```

# The scale_*_* function

When you set the symbology with the `aes()` function, you can customize 
the legend (and the details of the display), using a corresponding
`scale_*_*` function (I'm using stars there to represent a fill-in-the-blank).

For example, since I used the `aes()` function to set the point shape to vary by 
airport type, I can use `scale_shape()` to customize that legend. I can set
the legend title using `name = ` and I can set the labels for the individual 
legend elements using `labels = `.

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  scale_shape(name = "Airport type",
              labels = c("Major airport",
                         "Mid-sized airport")) +
  theme_void()
```

Alternatively, I can use `scale_shape_manual` if I also want to change the specific shapes that
are on the legend _and_ on the map. `16` refers to a solid dot and `1` refers to an open dot 
([for more on point symbology options, see this page](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-layer-symbology.html){target="_blank"}).

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  scale_shape_manual(name = "Airport type",
              labels = c("Major airport",
                         "Mid-sized airport"),
              values = c(16, 1)) +
  theme_void()
```

Likewise, I can use `scale_fill_continuous()` to set the title and labels for the color legend.
Since this is a continuous variable, I'll need to specify which values on the color ramp
will be labeled using `breaks = ` in addition to how they will be labeled.

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  scale_shape_manual(name = "Airport type",
              labels = c("Major airport",
                         "Mid-sized airport"),
              values = c(16, 1)) +
  scale_fill_continuous(name = "Estimated GDP",
                        breaks = c(100000,
                                   200000,
                                   300000,
                                   400000),
              labels = c("$100 Billion",
                         "$200 Billion",
                         "$300 Billion",
                         "$400 Billion")) +
  theme_void()
```


There are some shortcuts to defining breaks and labels. The `seq()` function
will create a set of sequential numbers for you, so save you having to 
type them all out. Here is how I would specify that I want breaks from 100,000 to 
400,000, increasing in units of 100,000.

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  scale_shape_manual(name = "Airport type",
              labels = c("Major airport",
                         "Mid-sized airport"),
              values = c(16, 1)) +
  scale_fill_continuous(name = "Estimated GDP",
                        breaks = seq(100000, 400000,
                                     by=100000),
              labels = c("$100 Billion",
                         "$200 Billion",
                         "$300 Billion",
                         "$400 Billion")) +
  theme_void()
```

If I set the break values as breaks = my_breaks <- seq(...), that will 
temporarily store those break values and I can use them to set up my labels.

The `paste0()` function is handy for setting up labels. In the example below,
I'm specifying that the labels should start with a "$", followed by the break 
value divided by 1000 (since the variable reports the value in millions), 
followed by a space and the word "Billion."

```{r}
ggplot() +
  geom_sf(data = countries,
          aes(fill = gdp_md_est)) +
  geom_sf(data = airports, 
          aes(shape = type)) +
  scale_shape_manual(name = "Airport type",
              labels = c("Major airport",
                         "Mid-sized airport"),
              values = c(16, 1)) +
  scale_fill_continuous(name = "Estimated GDP",
                        breaks = my_breaks <- seq(100000, 400000,
                                                  by=100000),
              labels = paste0("$", my_breaks / 1000, " Billion")) +
  theme_void()
```

You can also customize the color ramp using variations on `scale_fill_gradient()`
or `scale_fill_viridis_c()`. 
More on that in the [fun with color palettes skill](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-colors.html){target="_blank"}.

To further customize the appearance of the legend, refer to [customizing ggplot themes](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-custom-themes.html){target="_blank"}.
