---
title: "Fun with colors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are a lot of great tools for defining and applying color palettes for 
use with ggplot in R.

# Required packages

These example uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(MetBrewer)
library(RColorBrewer)
library(ggthemes)
```


# Loading some sample data

I'll start with data on counties in the contiguous United States. I've joined it 
with a spreadsheet of Urban-Rural classifications from the
[United States Department of Agriculture](https://www.ers.usda.gov/data-products/rural-urban-continuum-codes/){target="_blank"}
so that I can demonstrate with a categorical variable as well. Here are the first few rows:

```{r, message=FALSE, echo=FALSE, results='hide'}
vars <- c(pop = "P002001",
              ua_pop = "P002003")

county_data <- tidycensus::get_decennial(geography = "county", 
                          variables = "P002003",
                          summary_var = "P002001") %>%
  mutate(pct_ua = value / summary_value)

county_classes <- here::here("a2-files",
                       "Examples",
                       "ruralurbancodes2013.xls") %>%
  readxl::read_xls() %>%
  rename(GEOID = FIPS)
  

counties <- tigris::counties(year = 2010) %>%
  filter(STATEFP10 != "60" &
           STATEFP10 != "66" &
           STATEFP10 != "69" &
           STATEFP10 != "72" &
           STATEFP10 != "78" &
           STATEFP10 != "02" &
           STATEFP10 != "15") %>%
  rename(GEOID = GEOID10) %>%
  left_join(county_classes, by = "GEOID") %>%
  left_join(county_data, by = "GEOID") %>%
  select(NAMELSAD10,
         Description,
         pct_ua)
```

```{r, echo=FALSE}
knitr::kable(head(counties, 25),
             digits = 2)
```


# Default map of continuous variables

Here is a cholorpleth map showing the variation in the percentage of each 
county's population that lives in a census tract that is classified as
being in an urbanized area. By default, continuous variables are represented
by shades of blue.

```{r}
continuous_map <- ggplot(counties) +
  geom_sf(aes(fill = pct_ua)) +
  coord_sf(crs = 5070) +
  theme_map()

continuous_map
```

Here is a cholorpleth map showing the urban-rural classification categories. The 
default for categorical variables is to select colors that are evenly spaced
around a color ramp. 

```{r, fig.height=7}
categorical_map <- ggplot(counties) +
  geom_sf(aes(fill = Description)) +
  coord_sf(crs = 5070) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.direction = "vertical")

categorical_map
```

# Referencing colors

Before you can get too far into customizing colors on R plots, it can be useful
to know how to refer to specific colors in your code. There are hundreds of different colors you can refer to by name in R, as shown below. In addition to 
these, you can refer to a shade of gray with the with the word `gray` (or 
`grey`) followed by a number ranging from 0 to 100 (e.g. `gray30` or `gray80`). 
If you leave off the number (just `gray`), you'll get the same thing as `gray75`.

```{r, fig.height=20, echo=FALSE, out.width="100%"}
# No margin around chart
par(mar=c(0,0,0,0))

no_gray_colors <- c(colors()[0:151], colors()[254:259], colors()[357:657])


# Empty chart
plot(0, 0, 
     type = "n", 
     xlim = c(0, 1), 
     ylim = c(0, 1), 
     axes = FALSE, 
     xlab = "", 
     ylab = "")

# Settings
line <- 92
col <- 5

# Add color background
rect(  
  rep((0:(col - 1)/col),line) ,  
  sort(rep((0:(line - 1)/line),col),decreasing=T),   
  rep((1:col/col),line) , 
  sort(rep((1:line/line),col),decreasing=T),  
  border = "white" , 
  col=no_gray_colors[seq(1,line*col)])

# Color names
text(  
  rep((0:(col - 1)/col),line)+(1/(2*col)) ,  
  sort(rep((0:(line - 1)/line),col),decreasing=T)+1/(2*line), 
  no_gray_colors[seq(1,line*col)]  , 
  cex=1)
```

You can also refer to colors by their RGB (red-green-blue) hex codes. These will
be six-digit strings preceded by a `#`. There are a number of web-based tools for 
finding a color's hex code, like [this one](https://www.w3schools.com/colors/colors_picker.asp){target="_blank"}.

For a categorical variable, I can define a color scale using the 
`scale_fill_manual()` function (or `scale_color_manual()` for points and lines).
Again, I can reference colors by their names:

```{r, fig.height=7}
categorical_map +
  scale_fill_manual(values = c("white",
                               "gray",
                               "black",
                               "red",
                               "orange",
                               "yellow",
                               "green",
                               "blue",
                               "purple"))
```

Or by hex codes: 

```{r, fig.height=7}
categorical_map +
  scale_fill_manual(values = c("#ffffff",
                               "#999999",
                               "#000000",
                               "#ff0000",
                               "#ff9900",
                               "#ffff33",
                               "#00ff00",
                               "#0000ff",
                               "#9933ff"))
```

For a continuous variable, I can set up a gradient. `scale_fill_gradient()`
will let you set up a gradient from one color to another.

```{r}
continuous_map +
  scale_fill_gradient(low = "blue", high = "red")
```

`scale_fill_gradient2` will let you set up a gradient from one color, through
a midpoint, to another color. Helpfully, you can specify the value of your
midpoint if you don't want it to be the middle of the range. Here are three plots
of the same variable with different values for the midpoint on the color ramp.

This one places the midpoint in the middle of the range.

```{r}
continuous_map +
  scale_fill_gradient2(low = "green", mid = "yellow", high = "red",
                       midpoint = 0.5)
```

This one has a midpoint in the higher part of the range.

```{r}
continuous_map +
  scale_fill_gradient2(low = "green", mid = "yellow", high = "red",
                       midpoint = 0.75)
```

This one has a midpoint in the lower part of the range.

```{r}
continuous_map +
  scale_fill_gradient2(low = "green", mid = "yellow", high = "red",
                       midpoint = 0.25)
```

`scale_fill_gradientn` will let you create a color ramp that includes more 
than three colors.

```{r}
continuous_map +
  scale_fill_gradientn(colors = c("white",
                               "gray",
                               "black",
                               "red",
                               "orange",
                               "yellow",
                               "green",
                               "blue",
                               "purple"))
```

# Predefined color palettes

Manually defining the colors in a palette can be tedious, and you'll often 
end up with ugly color palettes. Luckily, there are several predefined color palettes you can use.

## Viridis

The [viridis](https://bids.github.io/colormap/){target="_blank"} color ramps are meant to be colorblind friendly and to reproduce reasonbly well in grayscale (for example, if someone prints your graphic then photocopies it with a black-and-white printer). There are 4 options.

Option A:

```{r}
continuous_map +
  scale_fill_viridis_c(option = "A")
```

Option B:

```{r}
continuous_map +
  scale_fill_viridis_c(option = "B")
```

Option C:

```{r}
continuous_map +
  scale_fill_viridis_c(option = "C")
```

Option D (this is the default if you don't specify an option):

```{r}
continuous_map +
  scale_fill_viridis_c(option = "D")
```

You can reverse the color ramp values by setting `begin = 1, end = 0`.

```{r}
continuous_map +
  scale_fill_viridis_c(option = "D", 
                       begin = 1, end = 0)
```

You can also use the viridis color palette for discrete variables:

```{r, fig.height=7}
categorical_map +
  scale_fill_viridis_d(option = "A")
```

## Color brewer

The `RColorBrewer` package offers a bunch of pre-defined palettes that can be useful for continuous and categorical data. You can view the options with the function `display.brewer.all()

```{r, fig.height=8}
display.brewer.all()
```

The top group of palettes you see above are sequential palettes, which range from dark to light colors.

To use a Color Brewer palette, you first set up the palette with the 
`brewer.pal()` function, indicating the number of colors you want and the
palette to draw them from.

```{r}
cat_palette <- brewer.pal(9, "Pastel1")
cont_palette <- brewer.pal(5, "Spectral")
```

Then you can apply that palette using `scale_fill_manual()` for categorical 
variables.

```{r, fig.height=7}
categorical_map +
  scale_fill_manual(values = cat_palette)
```

Or `scale_fill_gradientn()` for categorical variables.

```{r}
continuous_map +
  scale_fill_gradientn(colors = cont_palette)
```

As with the viridis functions, you can reverse the directions of a color ramp. 
You can do this by calling the `rev()` fuction on the color ramp you've 
defined.

```{r}
continuous_map +
  scale_fill_gradientn(colors = rev(cont_palette))
```

# Themed palettes

There are some other fun package that define color palettes, including 
[wesanderson](https://github.com/karthik/wesanderson){target="_blank"} (colors inspired by Wes Anderson 
films), [PNWColors](https://jakelawlor.github.io/portfolio/pnwcolors/){target="_blank"} (colors inspired by 
the package author's photos of the Pacific Northwest), 
[MetBrewer](https://github.com/BlakeRMills/MetBrewer/tree/main){target="_blank"} (inspired by works at the 
Metropolitan Museum of Art), and [MexBrewer](https://paezha.github.io/MexBrewer/){target="_blank"} (inspired 
by works of Mexican muralists). These four packages are structured similarly. Some of the available palettes 
don't work well for continuous variables (the order of the colors isn't intuitive), but some do. Spend
some time experimenting.

```{r}
degas_contin <- met.brewer(name = "Degas", n = 5, type = "continuous")

continuous_map +
  scale_fill_gradientn(colors = degas_contin)
```

```{r, fig.height=7}
Redon_cat <- met.brewer(name = "Redon", n = 9, type = "discrete")

categorical_map +
  scale_fill_manual(values = Redon_cat)
```
