---
title: "Vary colors for leaflet features based on a variable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can vary the colors of lines, points, and polygons based on the 
values of the variables in those datasets.

# Required packages

These examples uses the following packages:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(leaflet)
library(sf)
library(tidycensus)
library(htmltools)
```

# Loading some sample data

I'll load some data with census tracts in San Luis Obispo County, California:

```{r, message=FALSE, results='hide'}
tracts <- get_acs(geography = "tract", 
                  variables = c(median_income = "B06011_001"), 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) %>%
  st_transform("WGS84")
```

And display them on a Leaflet map. 

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              fillColor = "springgreen",
              fillOpacity = 0.3,
              color = "darkgreen",
              weight = 1)
```

# Define a color palette

I might want to vary the color of the tracts based on median income. First, I'll 
need to define a color ramp. You can define a color ramp using the `colorNumeric()` 
function. This function has two required arguments. 

* `palette` can be the name of a 
viridis palette (`"viridis"`, `"magma"`, `"inferno"`, or `"plasma"` - see more about 
[viridis palettes here](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-colors.html){target="_blank"}).
It could also be name of an RColorBrewer palette (see more about 
[RColorBrewer palettes here](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-colors.html){target="_blank"}).
* `domain` should be the variable that the color palette maps to (preceded by the
name of the data frame it's a part of).


```{r}
income_palette = colorNumeric(palette = "RdYlBu", 
                              domain = tracts$median_incomeE)
```

# Apply the color palette

Now I can use apply that palette to the layer on my leaflet map.

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = tracts,
              color = "black",
              fillOpacity = 0.7,
              weight = 0.3,
              fillColor = ~income_palette(median_incomeE))
```

Unlike ggplot, leaflet won't automatically add a legend to interpret the colors.
You'll need to [add that in yourself](https://c-voulgaris.github.io/VIS-2128-F22/a2-files/skill-leaflet-legend.html){target="_blank"}.


