---
title: "Trails, Transit, and Tracts"
output: 
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This map allows readers to explore the county trails of San Luis Obispo
County in relation to the the locations of transit stops and census-tract-level
median incomes. It shows that, for the most part, county trails are not
accessible by public transit.

The intended audience for this map is a resident of San Luis Obispo County
who might appreciate the ability to interactively explore the trails and transit
stops that serve the places where they live, work, and recreate.

```{r load-packages, include=FALSE}
library(tidyverse)
library(leaflet)
library(sf)
library(tidycensus)
library(here)
library(htmltools)
```

```{r load-data, include=FALSE}
trails <- here("a1-files",
               "County_Trails.geojson") %>%
  st_read() %>%
  mutate(label_text = paste(TrailName, 
                             "<br/>", 
                             formatC(Length, digits = 2),
                            LengthUnits))

rta_stops <- here("a1-files",
               "bus-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
    mutate(`Transit Provider` = "San Luis Obispo RTA") 

slo_stops <- here("a2-files",
                  "Examples",
                  "slo-stops.csv") %>%
  st_read(options=c("X_POSSIBLE_NAMES=stop_lon",
                               "Y_POSSIBLE_NAMES=stop_lat")) %>%
    mutate(`Transit Provider` = "SLO Transit") %>%
  select(colnames(rta_stops)) 

all_stops <- rbind(rta_stops, slo_stops)

tracts <- get_acs(geography = "tract", 
                  variables = "B06011_001", 
                  output = "wide", 
                  state = "CA",
                  county = "San Luis Obispo",
                  geometry = TRUE) %>%
  st_transform("WGS84") %>%
  mutate(label_text = paste0(NAME, 
                             "<br/><b>Estimated median income:</b> $", 
                             prettyNum(B06011_001E, big.mark = ",")))
```

```{r, echo=FALSE}
income_pal <- colorNumeric(palette = "RdYlBu", domain = tracts$B06011_001E, na.color = NA)
transit_pal <- colorFactor(palette = c("darkblue", "darkgreen"), domain = all_stops$`Transit Provider`)

leaflet(width = "100%") %>%
  addProviderTiles(providers$Stamen.TonerLite,
                   options = providerTileOptions(minZoom = 9)) %>%
  setView(lng = -120.6592, lat = 35.28190, zoom = 11) %>%
  addPolygons(data = tracts,
              color = "black",
              weight = 0.3,
              fillColor = ~income_pal(B06011_001E),
              fillOpacity = 0.5,
              popup = ~label_text,
              highlightOptions = highlightOptions(fillOpacity = 0.7)) %>%
  addCircleMarkers(data = all_stops,
             lat = ~stop_lat,
             lng = ~stop_lon,
             radius = 4,
             stroke = FALSE,
             fillOpacity = 0.7,
             fillColor = ~transit_pal(`Transit Provider`),
             popup = ~stop_name) %>%
  addPolylines(data = trails,
               color = "#873e23",
               weight = 2,
               opacity = 1,
               popup = ~label_text,
               highlightOptions = highlightOptions(weight = 4)) %>%
  addLegend(data = tracts,
            "bottomleft", pal = income_pal, values = ~B06011_001E,
            title = "Estimated Median Income",
    labFormat = labelFormat(prefix = "$"),
    opacity = 0.3) %>%
  addLegend(data = all_stops,
            "bottomright", 
            pal = transit_pal, 
            values = ~`Transit Provider`,
            title = "Transit provider",
    opacity = 0.7)
```

#### *Date Sources:*

* Estimated median income from the 2019 American Community Survey (5-year sample), accessed via 
[tidycensus](https://CRAN.R-project.org/package=tidycensus){target="_blank"} package
* Trail data from [San Luis Opispo County Open Data](https://opendata.slocounty.ca.gov/datasets/county-trails/explore?location=35.284761%2C-120.639250%2C11.00){target="_blank"}
* SLO Transit stop data from the [SLO Transit GTFS feed](https://app.mecatran.com/urb/ws/feed/c2l0ZT1zbG90cmFuc2l0O2NsaWVudD1zZWxmO2V4cGlyZT07dHlwZT1ndGZzO2tleT0zZTMwMzM1OTRiMTE2NzA0N2IxNjQwNjA0ZjQwMGMzMzdiM2E1MTQ0){target="_blank"}
* RTA transit stop data from the [RTA GTFS feed](http://slo.connexionz.net/rtt/public/resource/gtfs.zip){target="_blank"}

#### *Acknowldgments:*

My classmate, Michael Lidwin, helped me with this assignment by showing me how to set the initial zoom level.