---
title: "Assignment 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

# Deliverable

Submit a map with at least two layers: a raster layer and a vector
layer. The raster layer should be an image of a map
from the Harvard Map Collection (go to the Map 
Collection at the Pusey library in Harvard Yard and a librarian will help you find
and digitize a map) or an image of a historic map from an online source. 
The other should be one of the layers you described in 
Assignment 1 (you can use more than one layer from Assignment 1 if you 
want to, but you only need to use one). 

Your map should be attractive and informative (inset maps, north arrows, and 
scale bars are very helpful in making a map informative, but I'm not 
dogmatic about including them if your audience doesn't need them). It should 
clearly indicate the source of each layer in your map. It should
also include a couple of sentences describing the audience for your map and
the argument you are making or the message you are conveying to them with
your map.

On your submission, list the name(s) of at least one student who was 
especially helpful you you in completing this assignment, and describe 
how they helped.

# Relevant Skills

You are unlikely to use all of these skills in this assignment, but all 
of them are relevant and you will probably use some of them.

* [Georeferencing a map image (AllMaps)](https://harvardmapcollection.github.io/tutorials/allmaps/georeference/){target="_blank"}
* [Georeferencing a map image (ArcGIS)](https://harvard.zoom.us/rec/share/TruZkY4n-1zoylfrUpbc8MBSSIrW3j07DTqnCHAtLqCUYrl-0ZcM8HCc3K5lGTm2.seGOOjpp0slak0x3?startTime=1602621784000){target="_blank"}
* [Georeferencing a map image (QGIS)](https://harvard.zoom.us/rec/play/_pwa91L5_JufyagpsybCG1Cb2yZ-vZeFWpEIss-SOui3dGSqUNjuJrEF_BFC1Ju3WbgUMOe1EKOmAnqW.HkmVHJXQGNTVxyT2?startTime=1602622287000&_x_zm_rtaid=RCRvcGukTr-LeAWdaueuWg.1660310247547.f660848fc828b97cf7ead655d36fbc62&_x_zm_rhtaid=879){target="_blank"}
* [Labeling a vector layer (QGIS)](https://harvard.zoom.us/rec/play/MtL_kTlUXVzf-BiB4Sv9wSPVZlYrsOxoV2tuqXNbF_D4E_K-URdGmdpqDxfFUiNAWDjbL7Qf8fWxShRH.xNeKdoociZORkh8t?autoplay=true&startTime=1663009291000){target="_blank"}
* [Load an image as a raster layer (R - ggplot)](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/skill-load-image-ggplot.html){target="_blank"}
* [Load map tiles from AllMaps as a leaflet background (R - leaflet)](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/skill-load-leaflet-georef-tiles.html){target="_blank"}
* [Interactive layer toggle (R - leaflet)](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/skill-leaflet-layer-ctrl.html){target="_blank"}

# Other resources

* [Library of Congress maps](https://www.loc.gov/collections/?fa=original-format%3Amap%7Cpartof%3Ageography+and+map+division&c=160){target="_blank"}
* [All Harvard Map Collection tutorials](https://harvardmapcollection.github.io/tutorials/){target="_blank"}
* [Harvard Library Digital Collection](https://library.harvard.edu/digital-collections){target="_blank"}

# Examples

Each of the examples below fully meets the requirements for this assignment.

* [Example 1](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/Examples/boston-sea-level.html){target="_blank"}
* [Example 2](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/Examples/africa-georef.pdf){target="_blank"}
* [Example 3](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/Examples/Shanghai.pdf){target="_blank"}
* [Example 4](https://c-voulgaris.github.io/VIS-2128-F22/a3-files/Examples/IndianapolisGeoreference.pdf){target="_blank"}

# Grading rubric

```{r, echo=FALSE}
rubric <- tibble(Item = c("Name of a helpful classmate",
                          "How your classmate helped",
                          "Inclusion of a georeferenced map image",
                          "Inclusion of a digital layer",
                          "The map is informative",
                          "The map is attractive",
                          "Source attribution",
                          "Description of audience",
                          "Description of argument/message",
                          "Total"),
                 `No credit` = c(rep("Missing", 9), ""),
                 `Half credit` = c("Not adequately described",
                                   "Not adequately described",
                                   "",
                                   "",
                                   "Minimal apparent effort",
                                   "Minimal apparent effort",
                                   "Not adequately described",
                                   "Not adequately described",
                                   "Not adequately described", 
                                   ""),
                 `Full credit` = c("Adequately described",
                                   "Adequately described",
                                   "Included",
                                   "Included",
                                   "Evident effort",
                                   "Evident effort",
                                   "Adequately described",
                                   "Adequately described",
                                   "Adequately described", 
                                   ""),
                 `Possible points` = c(2, 2, 4, 4, 3, 3, 1, 3, 3, 25))

kable(rubric)
```
