---
title: "Load georeferenced image as a leaflet basemap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If you georeference your image using [AllMaps](https://harvardmapcollection.github.io/tutorials/allmaps/georeference/){target="_blank"},
you can use it as a background image for a leaflet map. 

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(here)
```

# Add maps tiles from AllMaps

If you follow the process in the AllMaps tutorial, you'll end up with a link to 
a url template. You can use this as an argument in the the `addTiles()` function
for a leaflet map.

```{r}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15) 
```

# Add a vector layer

Now I'll load a layer of polygons that shows the areas of Boston that would be underwater at high tide
in the event of a 36-inch rise in sea level (which could happen in the 2070s).

```{r, message=FALSE, results='hide'}
hi_tide_36_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation.geojson") %>%
  st_read()
```

And I can display that layer on the top of the map tiles from my historic map.

```{r}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  addPolygons(data = hi_tide_36_inch,
              weight = 1) %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15) 
```




