---
title: "Adding a layer control to a leaflet map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can use a layer control in leaflet to allow your users
to interactively turn layers on and off. The documentation
for the `leaflet` package has some 
[explanation of layer controls here](https://rstudio.github.io/leaflet/showhide.html){target="_blank"}.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(here)
```

# Sample map

```{r, include=FALSE}
hi_tide_36_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation.geojson") %>%
  st_read()

hi_tide_21_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation-21in.geojson") %>%
  st_read()

hi_tide_9_inch <- here("a3-files",
                        "Examples",
                        "Climate_Ready_Boston_Sea_Level_Rise_Inundation-9in.geojson") %>%
  st_read()

bikes <- here("a3-files",
              "Examples",
              "Blue_Bike_Stations.geojson") %>%
  st_read()
```


Here is a historic map of Boston with polygon layers showing
9-inch (dark blue) and 36-inch (light blue) sea-level rise 
scenarios, along with black circles for the locations of bike 
share stations.

```{r}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  addPolygons(data = hi_tide_36_inch,
              weight = 1,
              fillColor = "cornflowerblue",
              fillOpacity = 0.9) %>%
  addPolygons(data = hi_tide_9_inch,
              weight = 1,
              fillColor = "darkblue",
              fillOpacity = 0.9) %>%
  addCircleMarkers(data = bikes,
                   color = "black",
                   radius = 2,
                   opacity = 1) %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15)
```

# Toggle between base groups

I can use a base group to allow the user to toggle between the
9-inch and 36-inch sea-level rise scenarios.

I need to define a group name for each layer I want to toggle 
among by setting the `group = ` parameter in the `addPolygons()`
function. The name of the group should be the name that you
want to appear on the layer control (which basically functions 
as a legend).

Then, I need to list those as base groups in the 
`addLayersControl()` function. I also need to set
`options = layersControlOptions(collapsed = FALSE)` so that
the layer controls are initially visible.

```{r}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  addPolygons(data = hi_tide_36_inch,
              weight = 1,
              fillColor = "cornflowerblue",
              fillOpacity = 0.9,
              group = "36-inch sea level rise") %>%
  addPolygons(data = hi_tide_9_inch,
              weight = 1,
              fillColor = "darkblue",
              fillOpacity = 0.9,
              group = "9-inch sea level rise") %>%
  addCircleMarkers(data = bikes,
                   color = "black",
                   radius = 2,
                   opacity = 1) %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15) %>%
  addLayersControl(baseGroups = c("9-inch sea level rise",
                                  "36-inch sea level rise"),
    options = layersControlOptions(collapsed = FALSE)) 
```

# Turn overlay groups on and off

I can use overlay groups to allow the user to turn the layer of
bike stations on and off.

Again, I need to name the group within the `addCircleMarkers()`
function. Then I list the name of that group as an overlay 
group.

```{r}
leaflet() %>%
  addTiles(urlTemplate = "https://allmaps.xyz/maps/ad0210f48b2636b9/{z}/{x}/{y}.png") %>%
  addPolygons(data = hi_tide_36_inch,
              weight = 1,
              fillColor = "cornflowerblue",
              fillOpacity = 0.9,
              group = "36-inch sea level rise") %>%
  addPolygons(data = hi_tide_9_inch,
              weight = 1,
              fillColor = "darkblue",
              fillOpacity = 0.9,
              group = "9-inch sea level rise") %>%
  addCircleMarkers(data = bikes,
                   color = "black",
                   radius = 2,
                   opacity = 1,
                   group = "Blue Bike Stations") %>%
  setView(lng = -71.052164, lat = 42.360081, zoom = 15) %>%
  addLayersControl(baseGroups = c("9-inch sea level rise",
                                  "36-inch sea level rise"),
                   overlayGroups = c("Blue Bike Stations"),
    options = layersControlOptions(collapsed = FALSE)) 
```
