---
title: "Adding georefernced layer to a map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Once you've georeferenced an image, you can add it to a ggplot map as a
raster layer.

# Required packages

This example uses the following packages

```{r, results='hide', message=FALSE}
library(tidyverse)
library(sf)
library(here)
library(raster)
library(ggspatial)
library(rnaturalearth)
```

# Load the image as a raster object

First you will need to use the `stack()` function to read the image file as a 
RasterStack.

```{r}
africa <- here("a3-files",
                 "large-image-files",
                 "africa_modified.tif") %>%
  stack()
```

# Adding the image to a ggplot

Then you can use the `layer_spatial()` function from the `ggspatial` package 
to add the raster layer to a ggplot.

```{r}
ggplot() +
  layer_spatial(africa)
```

# Adding a vector layer

Now you can add some vector layers over the map image. Here, I'll add modern 
national borders

```{r}
countries <- ne_countries(continent = "Africa",
                          returnclass = "sf") 


ggplot() +
  layer_spatial(africa) +
  geom_sf(data = countries,
          fill = NA,
          color = "black",
          size = 1)
```

